---
title: "Clim_dist_fitness_models"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code for models relating climate distance to fitness (Table 2-3)

To do: Try the method with holding geo distance as the average on prob of rep 
Could try with median geo distance 

## Libraries
```{r}
library(tidyverse)
library(lmerTest) #mixed models
library(broom.mixed) #tidy mixed model results 
library(tidymodels) #model workflows 
library(multilevelmod)
library(performance) #for calculating rsq for glmmerMod objects
```

## Read in Fitness Data
```{r}
wl2_establishment <- read_csv("../Processed.Data/WL2_Establishment.csv")
first_yr_surv <- read_csv("../Processed.Data/WL2_Y1Surv.csv")

winter_surv <- read_csv("../Processed.Data/WL2_WinterSurv.csv")
wl2_surv_to_bud_y2 <- read_csv("../Processed.Data/WL2_Surv_to_Rep_Y2.csv")
wl2_fruits_y2 <- read_csv("../Processed.Data/WL2_Fruits_Y2.csv")

wl2_prob_rep <- read_csv("../Processed.Data/WL2_ProbFit.csv") %>% 
  rename(AvgGD_Recent_Wtr_Year=`AvgGD_Recent_Water Year`,
         AvgGD_Recent_GrwSsn=`AvgGD_Recent_Growth Season`,
         AvgGD_Historic_Wtr_Year=`AvgGD_Historic_Water Year`,
         AvgGD_Historic_GrwSsn=`AvgGD_Historic_Growth Season`)
wl2_tot_fecundity <- read_csv("../Processed.Data/WL2_TotalRepOutput.csv") %>% 
  rename(AvgGD_Recent_Wtr_Year=`AvgGD_Recent_Water Year`,
         AvgGD_Recent_GrwSsn=`AvgGD_Recent_Growth Season`,
         AvgGD_Historic_Wtr_Year=`AvgGD_Historic_Water Year`,
         AvgGD_Historic_GrwSsn=`AvgGD_Historic_Growth Season`)

```

## Read in Size Pre-Transplant 
```{r}
wl2_pretrans_size1 <- read_csv("../Raw.Data/WL2_DNA_Collection_Size_survey_combined20230703_corrected.csv")

wl2_pretrans_size2 <- read_csv("../Raw.Data/WL2_Extras_DNA_collection_size_survey_combined20230706_corrected.csv") %>% 
  filter(!is.na(`height (cm)`)) #to get rid of genotypes that were measured on the other data sheet (NAs on this sheet)

wl2_pretrans_size_all <- bind_rows(wl2_pretrans_size1, wl2_pretrans_size2) %>%
  rename(pop=Pop, height.cm = `height (cm)`, long.leaf.cm=`longest leaf (cm)`) %>% 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  select(-DNA)
head(wl2_pretrans_size_all)
```

## Scale Predictors, Data Transformations, and Add Pre-transplant size 
```{r}
wl2_establishment_scaled <- wl2_establishment %>% 
  left_join(wl2_pretrans_size_all) %>% 
  mutate(across(c("GD_Recent_Wtr_Year_2023", "GD_Historic_Wtr_Year_2023",
              "GD_Recent_GrwSsn_2023", "GD_Historic_GrwSsn_2023",
              "TempDist_Historic_GrwSsn_2023", "TempDist_Historic_Wtr_Year_2023",
              "TempDist_Recent_GrwSsn_2023", "TempDist_Recent_Wtr_Year_2023",
              "PPTDist_Historic_GrwSsn_2023", "PPTDist_Historic_Wtr_Year_2023",
              "PPTDist_Recent_GrwSsn_2023", "PPTDist_Recent_Wtr_Year_2023",
              "Geographic_Dist", "height.cm"), scale, .names = "{.col}_scaled")) 

wl2_y1_surv_scaled <- first_yr_surv %>% 
  left_join(wl2_pretrans_size_all) %>% 
  mutate(across(c("GD_Recent_Wtr_Year_2023", "GD_Historic_Wtr_Year_2023",
              "GD_Recent_GrwSsn_2023", "GD_Historic_GrwSsn_2023",
              "TempDist_Historic_GrwSsn_2023", "TempDist_Historic_Wtr_Year_2023",
              "TempDist_Recent_GrwSsn_2023", "TempDist_Recent_Wtr_Year_2023",
              "PPTDist_Historic_GrwSsn_2023", "PPTDist_Historic_Wtr_Year_2023",
              "PPTDist_Recent_GrwSsn_2023", "PPTDist_Recent_Wtr_Year_2023",
              "Geographic_Dist", "height.cm"), scale, .names = "{.col}_scaled")) 

wl2_winter_surv_scaled <- winter_surv %>% 
  left_join(wl2_pretrans_size_all) %>% 
  filter(pop!="WR") %>% #remove pops with too low sample size 
  mutate(across(c("GD_Recent_Wtr_Year_2024", "GD_Historic_Wtr_Year_2024",
              "TempDist_Historic_Wtr_Year_2024","TempDist_Recent_Wtr_Year_2024",
              "PPTDist_Historic_Wtr_Year_2024", "PPTDist_Recent_Wtr_Year_2024",
              "Geographic_Dist", "height.cm"), scale, .names = "{.col}_scaled"))

wl2_surv_to_rep_y2_scaled <- wl2_surv_to_bud_y2 %>% 
  left_join(wl2_pretrans_size_all) %>% 
  filter(pop != "LV1", pop != "SQ1", pop != "WR") %>% #remove pops with too low sample size 
  mutate(across(c("GD_Recent_Wtr_Year_2024", "GD_Historic_Wtr_Year_2024",
              "GD_Recent_GrwSsn_2024", "GD_Historic_GrwSsn_2024",
              "TempDist_Historic_GrwSsn_2024", "TempDist_Historic_Wtr_Year_2024",
              "TempDist_Recent_GrwSsn_2024", "TempDist_Recent_Wtr_Year_2024",
              "PPTDist_Historic_GrwSsn_2024", "PPTDist_Historic_Wtr_Year_2024",
              "PPTDist_Recent_GrwSsn_2024", "PPTDist_Recent_Wtr_Year_2024",
              "Geographic_Dist", "height.cm"), scale, .names = "{.col}_scaled"))

wl2_fruits_y2_scaled <- wl2_fruits_y2 %>% 
  left_join(wl2_pretrans_size_all) %>% 
  filter(pop!="LV1", pop!="SQ1", pop!="WR") %>% #remove pops with too low sample size 
  mutate(logFruits=log(y2_fruits)) %>%  #transform fruit number with a log transformation 
  mutate(across(c("GD_Recent_Wtr_Year_2024", "GD_Historic_Wtr_Year_2024",
              "GD_Recent_GrwSsn_2024", "GD_Historic_GrwSsn_2024",
              "TempDist_Historic_GrwSsn_2024", "TempDist_Historic_Wtr_Year_2024",
              "TempDist_Recent_GrwSsn_2024", "TempDist_Recent_Wtr_Year_2024",
              "PPTDist_Historic_GrwSsn_2024", "PPTDist_Historic_Wtr_Year_2024",
              "PPTDist_Recent_GrwSsn_2024", "PPTDist_Recent_Wtr_Year_2024",
              "Geographic_Dist", "height.cm"), scale, .names = "{.col}_scaled"))
  
wl2_FrFlN_y2_scaled <- wl2_fruits_y2 %>% 
  left_join(wl2_pretrans_size_all) %>% 
  filter(pop!="LV1", pop!="SQ1", pop!="WR") %>% #remove pops with too low sample size 
  drop_na(FrFlN_y2) %>% 
  mutate(logFrFLs=log(FrFlN_y2)) %>% #transform fruit number with a log transformation 
  mutate(across(c("GD_Recent_Wtr_Year_2024", "GD_Historic_Wtr_Year_2024",
              "GD_Recent_GrwSsn_2024", "GD_Historic_GrwSsn_2024",
              "TempDist_Historic_GrwSsn_2024", "TempDist_Historic_Wtr_Year_2024",
              "TempDist_Recent_GrwSsn_2024", "TempDist_Recent_Wtr_Year_2024",
              "PPTDist_Historic_GrwSsn_2024", "PPTDist_Historic_Wtr_Year_2024",
              "PPTDist_Recent_GrwSsn_2024", "PPTDist_Recent_Wtr_Year_2024",
              "Geographic_Dist", "height.cm"), scale, .names = "{.col}_scaled"))

wl2_prob_fitness_scaled <- wl2_prob_rep %>% 
  left_join(wl2_pretrans_size_all) %>% 
  mutate(across(c("AvgGD_Historic_GrwSsn", "AvgGD_Historic_Wtr_Year",
              "AvgGD_Recent_GrwSsn", "AvgGD_Recent_Wtr_Year",
              "AvgTempDist_Historic_GrwSsn", "AvgTempDist_Historic_Wtr_Year",
              "AvgTempDist_Recent_GrwSsn", "AvgTempDist_Recent_Wtr_Year",
              "AvgPPTDist_Historic_GrwSsn", "AvgPPTDist_Historic_Wtr_Year",
              "AvgPPTDist_Recent_GrwSsn", "AvgPPTDist_Recent_Wtr_Year",
              "Geographic_Dist", "height.cm"), scale, .names = "{.col}_scaled")) 

wl2_rep_output_scaled <- wl2_tot_fecundity %>% 
  left_join(wl2_pretrans_size_all) %>% 
  mutate(logTotalFecundity=log(Total_Fecundity)) %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  mutate(across(c("AvgGD_Historic_GrwSsn", "AvgGD_Historic_Wtr_Year",
              "AvgGD_Recent_GrwSsn", "AvgGD_Recent_Wtr_Year",
              "AvgTempDist_Historic_GrwSsn", "AvgTempDist_Historic_Wtr_Year",
              "AvgTempDist_Recent_GrwSsn", "AvgTempDist_Recent_Wtr_Year",
              "AvgPPTDist_Historic_GrwSsn", "AvgPPTDist_Historic_Wtr_Year",
              "AvgPPTDist_Recent_GrwSsn", "AvgPPTDist_Recent_Wtr_Year",
              "Geographic_Dist", "height.cm"), scale, .names = "{.col}_scaled")) 
```

## Create dataframes for model predictions

### Geo and Water Year PPT dist fixed 
```{r}
geo_ppt_fixed_2023 <- wl2_establishment_scaled %>% 
  summarise(Geographic_Dist=mean(Geographic_Dist),
            Geographic_Dist_scaled=mean(Geographic_Dist_scaled),
            PPTDist_Recent_Wtr_Year_2023=mean(PPTDist_Recent_Wtr_Year_2023),
            PPTDist_Recent_Wtr_Year_2023_scaled=mean(PPTDist_Recent_Wtr_Year_2023_scaled))

#summary(wl2_establishment_scaled) #get range for tempdist 

pred_df_recent_wtr_yr_temp_2023 <- tibble(TempDist_Recent_Wtr_Year_2023 = seq(from = -3, to=10, length.out=46)) %>% 
  mutate(TempDist_Recent_Wtr_Year_2023_scaled = scale(TempDist_Recent_Wtr_Year_2023)) %>% 
  bind_cols(geo_ppt_fixed_2023) 

geo_ppt_fixed_2024 <- wl2_surv_to_rep_y2_scaled %>% 
  summarise(Geographic_Dist=mean(Geographic_Dist),
            Geographic_Dist_scaled=mean(Geographic_Dist_scaled),
            PPTDist_Recent_Wtr_Year_2024=mean(PPTDist_Recent_Wtr_Year_2024),
            PPTDist_Recent_Wtr_Year_2024_scaled=mean(PPTDist_Recent_Wtr_Year_2024_scaled))

#summary(wl2_surv_to_rep_y2_scaled) #get range for tempdist 

pred_df_recent_wtr_yr_temp_2024 <- tibble(TempDist_Recent_Wtr_Year_2024 = seq(from = -2, to=8, length.out=46)) %>% 
  mutate(TempDist_Recent_Wtr_Year_2024_scaled = scale(TempDist_Recent_Wtr_Year_2024)) %>% 
  bind_cols(geo_ppt_fixed_2024) 

geo_avgppt_fixed <- wl2_prob_fitness_scaled %>% 
  summarise(Geographic_Dist=mean(Geographic_Dist),
            Geographic_Dist_scaled=mean(Geographic_Dist_scaled),
            AvgPPTDist_Recent_Wtr_Year=mean(AvgPPTDist_Recent_Wtr_Year),
            AvgPPTDist_Recent_Wtr_Year_scaled=mean(AvgPPTDist_Recent_Wtr_Year_scaled))

#summary(wl2_prob_fitness_scaled) #get range for avgtempdist 

pred_df_recent_wtr_yr_avgtemp <- tibble(AvgTempDist_Recent_Wtr_Year = seq(from = -4, to=9, length.out=46)) %>% 
  mutate(AvgTempDist_Recent_Wtr_Year_scaled = scale(AvgTempDist_Recent_Wtr_Year)) %>% 
  bind_cols(geo_avgppt_fixed) 
```

### Geo and Water Year Temp dist fixed 
```{r}
geo_temp_fixed_2023 <- wl2_establishment_scaled %>% 
  summarise(Geographic_Dist=mean(Geographic_Dist),
            Geographic_Dist_scaled=mean(Geographic_Dist_scaled),
            TempDist_Recent_Wtr_Year_2023=mean(TempDist_Recent_Wtr_Year_2023),
            TempDist_Recent_Wtr_Year_2023_scaled=mean(TempDist_Recent_Wtr_Year_2023_scaled))

#summary(wl2_establishment_scaled) #get range for pptdist 

pred_df_recent_wtr_yr_ppt_2023 <- tibble(PPTDist_Recent_Wtr_Year_2023 = seq(from = -1903, to=-671, length.out=46)) %>% 
  mutate(PPTDist_Recent_Wtr_Year_2023_scaled = scale(PPTDist_Recent_Wtr_Year_2023)) %>% 
  bind_cols(geo_temp_fixed_2023) 
```

### Geo fixed for Clim Dist
```{r}
geo_fixed <- wl2_establishment_scaled %>% 
  summarise(Geographic_Dist=mean(Geographic_Dist),
            Geographic_Dist_scaled=mean(Geographic_Dist_scaled))

pred_df_gowers <- tibble(GD_Recent_Wtr_Year_2023 = seq(from = 0.35, to=0.70, length.out=46),
                         GD_Historic_GrwSsn_2023 = seq(from = 0.22, to=0.54, length.out=46),
                         GD_Recent_Wtr_Year_2024 = seq(from = 0.14, to=0.58, length.out=46),
                         GD_Recent_GrwSsn_2024 = seq(from = 0.15, to=0.50, length.out=46),
                         AvgGD_Recent_Wtr_Year = seq(from = 0.25, to=0.64, length.out=46),
                         AvgGD_Recent_GrwSsn = seq(from = 0.20, to=0.47, length.out=46)
                         ) %>% 
  mutate(across(c("GD_Recent_Wtr_Year_2023", "GD_Historic_GrwSsn_2023",
                  "GD_Recent_Wtr_Year_2024", "GD_Recent_GrwSsn_2024",
                  "AvgGD_Recent_Wtr_Year", "AvgGD_Recent_GrwSsn"),
                scale, .names = "{.col}_scaled")) %>% 
  bind_cols(geo_fixed)
```

### Geo and GrwSsn PPT dist fixed 
```{r}
geo_ppt_fixed_2023_grwssn <- wl2_establishment_scaled %>% 
  summarise(Geographic_Dist=mean(Geographic_Dist),
            Geographic_Dist_scaled=mean(Geographic_Dist_scaled),
            PPTDist_Recent_GrwSsn_2023=mean(PPTDist_Recent_GrwSsn_2023),
            PPTDist_Recent_GrwSsn_2023_scaled=mean(PPTDist_Recent_GrwSsn_2023_scaled))

#summary(wl2_establishment_scaled) #get range for tempdist 

pred_df_recent_grwssn_temp_2023 <- tibble(TempDist_Recent_GrwSsn_2023 = seq(from = -3, to=4, length.out=46)) %>% 
  mutate(TempDist_Recent_GrwSsn_2023_scaled = scale(TempDist_Recent_GrwSsn_2023)) %>% 
  bind_cols(geo_ppt_fixed_2023_grwssn) 

geo_avgppt_fixed_grwssn <- wl2_prob_fitness_scaled %>% 
  summarise(Geographic_Dist=mean(Geographic_Dist),
            Geographic_Dist_scaled=mean(Geographic_Dist_scaled),
            AvgPPTDist_Recent_GrwSsn=mean(AvgPPTDist_Recent_GrwSsn),
            AvgPPTDist_Recent_GrwSsn_scaled=mean(AvgPPTDist_Recent_GrwSsn_scaled))

pred_df_recent_grwssn_avgtemp <- tibble(AvgTempDist_Recent_GrwSsn = seq(from = -5, to=3, length.out=46)) %>% 
  mutate(AvgTempDist_Recent_GrwSsn_scaled = scale(AvgTempDist_Recent_GrwSsn)) %>% 
  bind_cols(geo_avgppt_fixed_grwssn) 
```

### Geo and GrwSsn Temp dist fixed
```{r}
geo_avgtemp_fixed_grwssn <- wl2_prob_fitness_scaled %>% 
  summarise(Geographic_Dist=mean(Geographic_Dist),
            Geographic_Dist_scaled=mean(Geographic_Dist_scaled),
            AvgTempDist_Recent_GrwSsn=mean(AvgTempDist_Recent_GrwSsn),
            AvgTempDist_Recent_GrwSsn_scaled=mean(AvgTempDist_Recent_GrwSsn_scaled))

pred_df_recent_grwssn_avgppt <- tibble(AvgPPTDist_Recent_GrwSsn = seq(from = -73, to=936, length.out=46)) %>% 
  mutate(AvgPPTDist_Recent_GrwSsn_scaled = scale(AvgPPTDist_Recent_GrwSsn)) %>% 
  bind_cols(geo_avgtemp_fixed_grwssn) 
```

### Gower's Clim Dist fixed 
Predictions based on Geo Distance 
```{r}
climdist_fixed <- wl2_establishment_scaled %>% 
  summarise(GD_Recent_Wtr_Year_2023=mean(GD_Recent_Wtr_Year_2023),
            GD_Recent_Wtr_Year_2023_scaled=mean(GD_Recent_Wtr_Year_2023_scaled))

pred_df_geo <- tibble(Geographic_Dist = seq(from = 0.35, to=0.70, length.out=46)) %>%
  mutate(Geographic_Dist_scaled = scale(Geographic_Dist)) %>% 
  bind_cols(climdist_fixed)
```

## Set Model Engine
```{r}
glmer.model_binomial <- 
  linear_reg() %>% 
  set_engine("glmer", family=binomial)
```

## Establishment Models
```{r}
#basic model workflow 
surv_wflow <- workflow() %>% 
  add_variables(outcomes = Establishment, predictors = c(pop, mf, block))
surv_fits <- tibble(wflow=list(
  pop.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ (1|pop) + (1|block))},
  
  pop.mf.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
surv_fits_wl2 <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_establishment_scaled))
surv_fits_wl2 %>% filter(name=="pop.mf.block") %>% pull(fit) #mf explains very little variation and get singularity warning

#Add in Geo and tmp/ppt dist
surv_GD_wflow_sub <- workflow() %>%
  add_variables(outcomes = Establishment, predictors = c(pop, block, height.cm_scaled, contains("Dist"))) 
surv_GD_fits_sub <- tibble(wflow=list(
  pop.block = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ height.cm_scaled + (1|pop) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ TempDist_Recent_Wtr_Year_2023_scaled + PPTDist_Recent_Wtr_Year_2023_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ TempDist_Historic_Wtr_Year_2023_scaled + PPTDist_Historic_Wtr_Year_2023_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  GS_Recent = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ TempDist_Recent_GrwSsn_2023_scaled + PPTDist_Recent_GrwSsn_2023_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  GS_Historical = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ TempDist_Historic_GrwSsn_2023_scaled + PPTDist_Historic_GrwSsn_2023_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  WY_Recent_height = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ TempDist_Recent_Wtr_Year_2023_scaled + PPTDist_Recent_Wtr_Year_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
  WY_Historical_height = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ TempDist_Historic_Wtr_Year_2023_scaled + PPTDist_Historic_Wtr_Year_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
  GS_Recent_height = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ TempDist_Recent_GrwSsn_2023_scaled + PPTDist_Recent_GrwSsn_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
  GS_Historical_height = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ TempDist_Historic_GrwSsn_2023_scaled + PPTDist_Historic_GrwSsn_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow) 
surv_GD_fits_wl2_sub <- surv_GD_fits_sub %>%
  mutate(fit = map(wflow, fit, data = wl2_establishment_scaled))

##show results:
surv_GD_fits_wl2_sub %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist") | term=="height.cm_scaled") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)

surv_GD_fits_wl2_sub %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>% 
  mutate(rsq=map(extracted_fit, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-wflow:-extracted_fit)
#conditional rsq = first row per model - based on fixed and random effects 
#and marginal rsq is the second row per model - based only on fixed effects

##predictions for figures:
est_sub_preds <- surv_GD_fits_wl2_sub %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="WY_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_recent_wtr_yr_temp_2023)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_recent_wtr_yr_temp_2023) %>% 
  select(-wflow, -fit, -extracted_fit) 
est_sub_preds$TempDist_Recent_Wtr_Year_2023_scaled <- as.vector(est_sub_preds$TempDist_Recent_Wtr_Year_2023_scaled)
write_csv(est_sub_preds, "../Processed.Data/Est_SubPreds_WYRecentTemp.csv")

est_sub_preds2 <- surv_GD_fits_wl2_sub %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="WY_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_recent_wtr_yr_ppt_2023)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_recent_wtr_yr_ppt_2023) %>% 
  select(-wflow, -fit, -extracted_fit) 
est_sub_preds2$PPTDist_Recent_Wtr_Year_2023_scaled <- as.vector(est_sub_preds2$PPTDist_Recent_Wtr_Year_2023_scaled)
write_csv(est_sub_preds2, "../Processed.Data/Est_SubPreds_WYRecentPPT.csv")

est_sub_preds3 <- surv_GD_fits_wl2_sub %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="GS_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_recent_grwssn_temp_2023)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_recent_grwssn_temp_2023) %>% 
  select(-wflow, -fit, -extracted_fit) 
est_sub_preds3$TempDist_Recent_GrwSsn_2023_scaled <- as.vector(est_sub_preds3$TempDist_Recent_GrwSsn_2023_scaled)
write_csv(est_sub_preds3, "../Processed.Data/Est_SubPreds_GrwSsnRecentTemp.csv")

#Add in Geo and Gowers clim dist
surv_GD_wflow <- workflow() %>%
  add_variables(outcomes = Establishment, predictors = c(pop, height.cm_scaled, block, contains("GD"), Geographic_Dist_scaled)) 
surv_GD_fits <- tibble(wflow=list(
  pop.block = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ height.cm_scaled + (1|pop) + (1|block))},
  
  WY_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GD_Recent_Wtr_Year_2023_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  WY_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GD_Historic_Wtr_Year_2023_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  GS_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GD_Recent_GrwSsn_2023_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  GS_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GD_Historic_GrwSsn_2023_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  WY_Recent_height = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GD_Recent_Wtr_Year_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
  WY_Historical_height = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GD_Historic_Wtr_Year_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
  GS_Recent_height = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GD_Recent_GrwSsn_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
  GS_Historical_height = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Establishment ~ GD_Historic_GrwSsn_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow) 
surv_GD_fits_wl2 <- surv_GD_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_establishment_scaled))

##show results:
surv_GD_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist_scaled" | term=="height.cm_scaled") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)

#surv_GD_fits_wl2 %>% mutate(glance=map(fit, glance)) %>% unnest(glance) %>% 
 # select(-wflow:-sigma) 

surv_GD_fits_wl2 %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>% 
  mutate(rsq=map(extracted_fit, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-wflow:-extracted_fit)
#conditional rsq = first row per model - based on fixed and random effects 
#and marginal rsq is the second row per model - based only on fixed effects

##predictions for figures:
est_GD_preds <- surv_GD_fits_wl2 %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="WY_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_gowers)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_gowers) %>% 
  select(-wflow, -fit, -extracted_fit) 
est_GD_preds$GD_Recent_Wtr_Year_2023_scaled <- as.vector(est_GD_preds$GD_Recent_Wtr_Year_2023_scaled)
est_GD_preds$GD_Historic_GrwSsn_2023_scaled <- as.vector(est_GD_preds$GD_Historic_GrwSsn_2023_scaled)
est_GD_preds$GD_Recent_Wtr_Year_2024_scaled <- as.vector(est_GD_preds$GD_Recent_Wtr_Year_2024_scaled)
est_GD_preds$GD_Recent_GrwSsn_2024_scaled <- as.vector(est_GD_preds$GD_Recent_GrwSsn_2024_scaled)
est_GD_preds$AvgGD_Recent_Wtr_Year_scaled <- as.vector(est_GD_preds$AvgGD_Recent_Wtr_Year_scaled)
est_GD_preds$AvgGD_Recent_GrwSsn_scaled <- as.vector(est_GD_preds$AvgGD_Recent_GrwSsn_scaled)
write_csv(est_GD_preds, "../Processed.Data/Est_GDPreds_WYRecent.csv")

est_GD_preds2 <- surv_GD_fits_wl2 %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="WY_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_geo)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_geo) %>% 
  select(-wflow, -fit, -extracted_fit) 
est_GD_preds2$Geographic_Dist_scaled <- as.vector(est_GD_preds2$Geographic_Dist_scaled)
write_csv(est_GD_preds2, "../Processed.Data/Est_GeoDist_Preds.csv")

est_GD_preds3 <- surv_GD_fits_wl2 %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="GS_Historical") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_gowers)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_gowers) %>% 
  select(-wflow, -fit, -extracted_fit) 
est_GD_preds3$GD_Recent_Wtr_Year_2023_scaled <- as.vector(est_GD_preds3$GD_Recent_Wtr_Year_2023_scaled)
est_GD_preds3$GD_Historic_GrwSsn_2023_scaled <- as.vector(est_GD_preds3$GD_Historic_GrwSsn_2023_scaled)
est_GD_preds3$GD_Recent_Wtr_Year_2024_scaled <- as.vector(est_GD_preds3$GD_Recent_Wtr_Year_2024_scaled)
est_GD_preds3$GD_Recent_GrwSsn_2024_scaled <- as.vector(est_GD_preds3$GD_Recent_GrwSsn_2024_scaled)
est_GD_preds3$AvgGD_Recent_Wtr_Year_scaled <- as.vector(est_GD_preds3$AvgGD_Recent_Wtr_Year_scaled)
est_GD_preds3$AvgGD_Recent_GrwSsn_scaled <- as.vector(est_GD_preds3$AvgGD_Recent_GrwSsn_scaled)
write_csv(est_GD_preds3, "../Processed.Data/Est_GDPreds_GrwSsnHist.csv")
```

## First Year Surv Models
```{r}
#basic model workflow 
surv_wflow <- workflow() %>% 
  add_variables(outcomes = Y1Survival, predictors = c(pop, mf, block))
surv_fits <- tibble(wflow=list(
  pop.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ (1|pop) + (1|block))},
  
  pop.mf.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
surv_fits_wl2 <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_y1_surv_scaled))
surv_fits_wl2 %>% filter(name=="pop.mf.block") %>% pull(fit) #no issues with mf 

#Add in Geo and tmp/ppt dist
surv_GD_wflow_sub <- workflow() %>%
  add_variables(outcomes = Y1Survival, predictors = c(pop, mf, height.cm_scaled, block, contains("Dist"))) 
surv_GD_fits_sub <- tibble(wflow=list(
  pop.block = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ TempDist_Recent_Wtr_Year_2023_scaled + PPTDist_Recent_Wtr_Year_2023_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ TempDist_Historic_Wtr_Year_2023_scaled + PPTDist_Historic_Wtr_Year_2023_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  GS_Recent = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ TempDist_Recent_GrwSsn_2023_scaled + PPTDist_Recent_GrwSsn_2023_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  GS_Historical = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ TempDist_Historic_GrwSsn_2023_scaled + PPTDist_Historic_GrwSsn_2023_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent_height = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ TempDist_Recent_Wtr_Year_2023_scaled + PPTDist_Recent_Wtr_Year_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical_height = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ TempDist_Historic_Wtr_Year_2023_scaled + PPTDist_Historic_Wtr_Year_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  GS_Recent_height = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ TempDist_Recent_GrwSsn_2023_scaled + PPTDist_Recent_GrwSsn_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  GS_Historical_height = {surv_GD_wflow_sub %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ TempDist_Historic_GrwSsn_2023_scaled + PPTDist_Historic_GrwSsn_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow) 
surv_GD_fits_wl2_sub <- surv_GD_fits_sub %>%
  mutate(fit = map(wflow, fit, data = wl2_y1_surv_scaled))

##show results
surv_GD_fits_wl2_sub %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist") | term=="height.cm_scaled") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)

surv_GD_fits_wl2_sub %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>% 
  mutate(rsq=map(extracted_fit, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-wflow:-extracted_fit)

##predictions for figures:
y1surv_sub_preds <- surv_GD_fits_wl2_sub %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="WY_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_recent_wtr_yr_temp_2023)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_recent_wtr_yr_temp_2023) %>% 
  select(-wflow, -fit, -extracted_fit) 
y1surv_sub_preds$TempDist_Recent_Wtr_Year_2023_scaled <- as.vector(y1surv_sub_preds$TempDist_Recent_Wtr_Year_2023_scaled)
write_csv(y1surv_sub_preds, "../Processed.Data/Y1Surv_SubPreds_WYRecentTemp.csv")

#Add in Geo and Gowers clim dist
surv_GD_wflow <- workflow() %>%
  add_variables(outcomes = Y1Survival, predictors = c(pop, mf, height.cm_scaled, block, contains("GD"), Geographic_Dist_scaled)) 
surv_GD_fits <- tibble(wflow=list(
  pop.block = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ GD_Recent_Wtr_Year_2023_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ GD_Historic_Wtr_Year_2023_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  GS_Recent = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ GD_Recent_GrwSsn_2023_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  GS_Historical = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ GD_Historic_GrwSsn_2023_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent_height = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ GD_Recent_Wtr_Year_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical_height = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ GD_Historic_Wtr_Year_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  GS_Recent_height = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ GD_Recent_GrwSsn_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  GS_Historical_height = {surv_GD_wflow %>% 
      add_model(glmer.model_binomial, formula = Y1Survival ~ GD_Historic_GrwSsn_2023_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow) 
surv_GD_fits_wl2 <- surv_GD_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_y1_surv_scaled))

##show results 
surv_GD_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist_scaled" | term=="height.cm_scaled") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)

surv_GD_fits_wl2 %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>% 
  mutate(rsq=map(extracted_fit, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-wflow:-extracted_fit)

##predictions for figures:
y1surv_GD_preds <- surv_GD_fits_wl2 %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="WY_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_gowers)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_gowers) %>% 
  select(-wflow, -fit, -extracted_fit) 
y1surv_GD_preds$GD_Recent_Wtr_Year_2023_scaled <- as.vector(y1surv_GD_preds$GD_Recent_Wtr_Year_2023_scaled)
y1surv_GD_preds$GD_Historic_GrwSsn_2023_scaled <- as.vector(y1surv_GD_preds$GD_Historic_GrwSsn_2023_scaled)
y1surv_GD_preds$GD_Recent_Wtr_Year_2024_scaled <- as.vector(y1surv_GD_preds$GD_Recent_Wtr_Year_2024_scaled)
y1surv_GD_preds$GD_Recent_GrwSsn_2024_scaled <- as.vector(y1surv_GD_preds$GD_Recent_GrwSsn_2024_scaled)
y1surv_GD_preds$AvgGD_Recent_Wtr_Year_scaled <- as.vector(y1surv_GD_preds$AvgGD_Recent_Wtr_Year_scaled)
y1surv_GD_preds$AvgGD_Recent_GrwSsn_scaled <- as.vector(y1surv_GD_preds$AvgGD_Recent_GrwSsn_scaled)
write_csv(y1surv_GD_preds, "../Processed.Data/Y1Surv_GDPreds_WYRecent.csv")

y1surv_GD_preds2 <- surv_GD_fits_wl2 %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="GS_Historical") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_gowers)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_gowers) %>% 
  select(-wflow, -fit, -extracted_fit) 
y1surv_GD_preds2$GD_Recent_Wtr_Year_2023_scaled <- as.vector(y1surv_GD_preds2$GD_Recent_Wtr_Year_2023_scaled)
y1surv_GD_preds2$GD_Historic_GrwSsn_2023_scaled <- as.vector(y1surv_GD_preds2$GD_Historic_GrwSsn_2023_scaled)
y1surv_GD_preds2$GD_Recent_Wtr_Year_2024_scaled <- as.vector(y1surv_GD_preds2$GD_Recent_Wtr_Year_2024_scaled)
y1surv_GD_preds2$GD_Recent_GrwSsn_2024_scaled <- as.vector(y1surv_GD_preds2$GD_Recent_GrwSsn_2024_scaled)
y1surv_GD_preds2$AvgGD_Recent_Wtr_Year_scaled <- as.vector(y1surv_GD_preds2$AvgGD_Recent_Wtr_Year_scaled)
y1surv_GD_preds2$AvgGD_Recent_GrwSsn_scaled <- as.vector(y1surv_GD_preds2$AvgGD_Recent_GrwSsn_scaled)
write_csv(y1surv_GD_preds2, "../Processed.Data/Y1Surv_GDPreds_GrwSsnHist.csv")
```

## Winter Surv Models
```{r}
#basic model workflow 
surv_wflow <- workflow() %>% 
  add_variables(outcomes = WinterSurv, predictors = c(pop, mf, block,  height.cm_scaled,))
surv_fits <- tibble(wflow=list(
  pop.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ height.cm_scaled + (1|pop) + (1|block))},
  
  pop.mf.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ height.cm_scaled + (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
surv_fits_wl2 <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_winter_surv_scaled))
surv_fits_wl2 %>% filter(name=="pop.mf.block") %>% pull(fit) #nothing wrong with the full model, so use that 

#Add in Geo and tmp/ppt dist
surv_GD_wflow_wl2_sub <- workflow() %>%
  add_variables(outcomes = WinterSurv, predictors = c(pop, mf,  height.cm_scaled, block, contains("Dist"))) 
surv_GD_fits_wl2_sub <- tibble(wflow=list(
  pop.mf.block = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ TempDist_Recent_Wtr_Year_2024_scaled + PPTDist_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ TempDist_Historic_Wtr_Year_2024_scaled + PPTDist_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent_height = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ TempDist_Recent_Wtr_Year_2024_scaled + PPTDist_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical_height = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ TempDist_Historic_Wtr_Year_2024_scaled + PPTDist_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_winter_surv_scaled))

##show results
surv_GD_fits_wl2_sub %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist") | term=="height.cm_scaled") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)

surv_GD_fits_wl2_sub %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>% 
  mutate(rsq=map(extracted_fit, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-wflow:-extracted_fit)

#Add in Geo and Gowers clim dist
surv_GD_wflow_wl2 <- workflow() %>%
  add_variables(outcomes = WinterSurv, predictors = c(pop, mf, height.cm_scaled,  block, contains("GD"), Geographic_Dist_scaled)) 
surv_GD_fits_wl2 <- tibble(wflow=list(
  pop.mf.block = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ GD_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ GD_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent_height = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ GD_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical_height = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = WinterSurv ~ GD_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))}
  
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_winter_surv_scaled))

##show results 
surv_GD_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist_scaled" | term=="height.cm_scaled") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)

surv_GD_fits_wl2 %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>% 
  mutate(rsq=map(extracted_fit, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-wflow:-extracted_fit)

##predictions for figures:
wintsurv_GD_preds <- surv_GD_fits_wl2 %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="WY_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_gowers)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_gowers) %>% 
  select(-wflow, -fit, -extracted_fit) 
wintsurv_GD_preds$GD_Recent_Wtr_Year_2023_scaled <- as.vector(wintsurv_GD_preds$GD_Recent_Wtr_Year_2023_scaled)
wintsurv_GD_preds$GD_Historic_GrwSsn_2023_scaled <- as.vector(wintsurv_GD_preds$GD_Historic_GrwSsn_2023_scaled)
wintsurv_GD_preds$GD_Recent_Wtr_Year_2024_scaled <- as.vector(wintsurv_GD_preds$GD_Recent_Wtr_Year_2024_scaled)
wintsurv_GD_preds$GD_Recent_GrwSsn_2024_scaled <- as.vector(wintsurv_GD_preds$GD_Recent_GrwSsn_2024_scaled)
wintsurv_GD_preds$AvgGD_Recent_Wtr_Year_scaled <- as.vector(wintsurv_GD_preds$AvgGD_Recent_Wtr_Year_scaled)
wintsurv_GD_preds$AvgGD_Recent_GrwSsn_scaled <- as.vector(wintsurv_GD_preds$AvgGD_Recent_GrwSsn_scaled)
write_csv(wintsurv_GD_preds, "../Processed.Data/WintSurv_GDPreds_WYRecent.csv")
```

## Surv to Buds Y2 Models
```{r}
#basic model workflow 
surv_wflow <- workflow() %>% 
  add_variables(outcomes = SurvtoRep_y2, predictors = c(pop, mf, block))
surv_fits <- tibble(wflow=list(
  pop.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ (1|pop) + (1|block))},
  
  pop.mf.block = {surv_wflow %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
surv_fits_wl2 <- surv_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_surv_to_rep_y2_scaled))
surv_fits_wl2 %>% filter(name=="pop.mf.block") %>% pull(fit) 
#convergence issues with mf included 

#Add in Geo and tmp/ppt dist
surv_GD_wflow_wl2_sub <- workflow() %>%
  add_variables(outcomes = SurvtoRep_y2, predictors = c(pop, mf, height.cm_scaled,  block, contains("Dist"))) 
surv_GD_fits_wl2_sub <- tibble(wflow=list(
  pop.block = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ height.cm_scaled + (1|pop) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ TempDist_Recent_Wtr_Year_2024_scaled + PPTDist_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ TempDist_Historic_Wtr_Year_2024_scaled + PPTDist_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  GS_Recent = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ TempDist_Recent_GrwSsn_2024_scaled + PPTDist_Recent_GrwSsn_2024_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  GS_Historical = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ TempDist_Historic_GrwSsn_2024_scaled + PPTDist_Historic_GrwSsn_2024_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  WY_Recent_height = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ TempDist_Recent_Wtr_Year_2024_scaled + PPTDist_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
 # WY_Historical_height = {surv_GD_wflow_wl2_sub %>% 
  #    add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ TempDist_Historic_Wtr_Year_2024_scaled + PPTDist_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))}
  
  GS_Recent_height = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ TempDist_Recent_GrwSsn_2024_scaled + PPTDist_Recent_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
  GS_Historical_height = {surv_GD_wflow_wl2_sub %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ TempDist_Historic_GrwSsn_2024_scaled + PPTDist_Historic_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_surv_to_rep_y2_scaled))
#mod_test <- glmer(SurvtoRep_y2 ~ TempDist_Historic_Wtr_Year_2024_scaled + PPTDist_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block), family = "binomial", data=wl2_surv_to_rep_y2_scaled) #convergence issues with WY_Historical_height model 

##show results 
surv_GD_fits_wl2_sub %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist") | term=="height.cm_scaled") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)

surv_GD_fits_wl2_sub %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>% 
  mutate(rsq=map(extracted_fit, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-wflow:-extracted_fit)

#Add in Geo and Gowers clim dist
surv_GD_wflow_wl2 <- workflow() %>%
  add_variables(outcomes = SurvtoRep_y2, predictors = c(pop, mf, height.cm_scaled,  block, contains("GD"), Geographic_Dist_scaled)) 
surv_GD_fits_wl2 <- tibble(wflow=list(
  pop.block = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ height.cm_scaled + (1|pop) + (1|block))},
  
  WY_Recent = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GD_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  WY_Historical = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GD_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  GS_Recent = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GD_Recent_GrwSsn_2024_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  GS_Historical = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GD_Historic_GrwSsn_2024_scaled + Geographic_Dist_scaled + (1|pop) + (1|block))},
  
  WY_Recent_height = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GD_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
  WY_Historical_height = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GD_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
  GS_Recent_height = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GD_Recent_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))},
  
  GS_Historical_height = {surv_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = SurvtoRep_y2 ~ GD_Historic_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_surv_to_rep_y2_scaled))

##show results:
surv_GD_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist_scaled" | term=="height.cm_scaled") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)

surv_GD_fits_wl2 %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>% 
  mutate(rsq=map(extracted_fit, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-wflow:-extracted_fit)

##predictions for figures:
repsurv2_GD_preds <- surv_GD_fits_wl2 %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="WY_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_gowers)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_gowers) %>% 
  select(-wflow, -fit, -extracted_fit) 
repsurv2_GD_preds$GD_Recent_Wtr_Year_2023_scaled <- as.vector(repsurv2_GD_preds$GD_Recent_Wtr_Year_2023_scaled)
repsurv2_GD_preds$GD_Historic_GrwSsn_2023_scaled <- as.vector(repsurv2_GD_preds$GD_Historic_GrwSsn_2023_scaled)
repsurv2_GD_preds$GD_Recent_Wtr_Year_2024_scaled <- as.vector(repsurv2_GD_preds$GD_Recent_Wtr_Year_2024_scaled)
repsurv2_GD_preds$GD_Recent_GrwSsn_2024_scaled <- as.vector(repsurv2_GD_preds$GD_Recent_GrwSsn_2024_scaled)
repsurv2_GD_preds$AvgGD_Recent_Wtr_Year_scaled <- as.vector(repsurv2_GD_preds$AvgGD_Recent_Wtr_Year_scaled)
repsurv2_GD_preds$AvgGD_Recent_GrwSsn_scaled <- as.vector(repsurv2_GD_preds$AvgGD_Recent_GrwSsn_scaled)
write_csv(repsurv2_GD_preds, "../Processed.Data/RepSurv2_GDPreds_WYRecent.csv")

repsurv2_GD_preds2 <- surv_GD_fits_wl2 %>%
  mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="GS_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_gowers)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_gowers) %>% 
  select(-wflow, -fit, -extracted_fit) 
repsurv2_GD_preds2$GD_Recent_Wtr_Year_2023_scaled <- as.vector(repsurv2_GD_preds2$GD_Recent_Wtr_Year_2023_scaled)
repsurv2_GD_preds2$GD_Historic_GrwSsn_2023_scaled <- as.vector(repsurv2_GD_preds2$GD_Historic_GrwSsn_2023_scaled)
repsurv2_GD_preds2$GD_Recent_Wtr_Year_2024_scaled <- as.vector(repsurv2_GD_preds2$GD_Recent_Wtr_Year_2024_scaled)
repsurv2_GD_preds2$GD_Recent_GrwSsn_2024_scaled <- as.vector(repsurv2_GD_preds2$GD_Recent_GrwSsn_2024_scaled)
repsurv2_GD_preds2$AvgGD_Recent_Wtr_Year_scaled <- as.vector(repsurv2_GD_preds2$AvgGD_Recent_Wtr_Year_scaled)
repsurv2_GD_preds2$AvgGD_Recent_GrwSsn_scaled <- as.vector(repsurv2_GD_preds2$AvgGD_Recent_GrwSsn_scaled)
write_csv(repsurv2_GD_preds2, "../Processed.Data/RepSurv2_GDPreds_GrwSsnRecent.csv")
```

## Fruit N Y2 Models
```{r}
#base models:
fruits_modelslog <- tribble(
  ~name,          ~f,
  "1_pop.block",        "logFruits ~ (1|pop) + (1|block)", 
  "2_pop.mf.block",     "logFruits ~  (1|pop/mf) + (1|block)"
)
#run the models 
fruits_modelslog <- fruits_modelslog %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_fruits_y2_scaled)), #run the models 
         glance = map(lmer, glance)) #glance at the model results
fruits_modelslog %>% mutate(tidy=map(lmer, tidy, effects = "ran_pars", scales = "vcov")) %>% unnest(tidy) %>% 
  select(name, group, estimate)

#Add in Geo and tmp/ppt dist
fruits_models_log_CD_GD_sub <- tribble(
  ~name,          ~f,
  "1_pop.block",      "logFruits ~  height.cm_scaled + (1|pop/mf) + (1|block)", 
  "2_WY_Recent",      "logFruits ~  TempDist_Recent_Wtr_Year_2024_scaled + PPTDist_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)",
  "3_WY_Historical",  "logFruits ~  TempDist_Historic_Wtr_Year_2024_scaled + PPTDist_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)",
  "4_GS_Recent",      "logFruits ~  TempDist_Recent_GrwSsn_2024_scaled + PPTDist_Recent_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)", 
  "5_GS_Historical",  "logFruits ~  TempDist_Historic_GrwSsn_2024_scaled + PPTDist_Historic_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)"
)
#run the models 
fruits_models_log_CD_GD_sub <- fruits_models_log_CD_GD_sub %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_fruits_y2_scaled)), #run the models 
         glance = map(lmer, glance)) #glance at the model results
#boundary (singular) fit: see help('isSingular') warning for all models with climate distance - pop explains 0 variance when you add clim distance 
##show results:
fruits_models_log_CD_GD_sub %>% mutate(tidy=map(lmer, tidy)) %>% unnest(tidy) %>%
  select(name, term:p.value) %>% 
  filter(str_detect(term, "Dist") | term=="height.cm_scaled") %>%
  drop_na(p.value)

fruits_models_log_CD_GD_sub %>% mutate(rsq=map(lmer, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-f:-glance)
# Can't compute random effect variances. Some variance components equal zero. --> NA conditional rsq
# Your model may suffer from singularity (see `?lme4::isSingular` and `?performance::check_singularity`).
#  Decrease the `tolerance` level to force the calculation of random effect variances, or impose
#  priors on your random effects parameters (using packages like `brms` or `glmmTMB`).

#Add in Geo and Gowers clim dist
fruits_models_log_CD_GD <- tribble(
  ~name,          ~f,
  "1_pop.block",      "logFruits ~  height.cm_scaled + (1|pop/mf) + (1|block)", 
  "2_WY_Recent",      "logFruits ~  GD_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)",
  "3_WY_Historical",  "logFruits ~  GD_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)",
  "4_GS_Recent",      "logFruits ~  GD_Recent_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)", 
  "5_GS_Historical",  "logFruits ~  GD_Historic_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)"
)
#run the models 
fruits_models_log_CD_GD <- fruits_models_log_CD_GD %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_fruits_y2_scaled)), #run the models 
         glance = map(lmer, glance)) #glance at the model results
##show results:
fruits_models_log_CD_GD %>% mutate(tidy=map(lmer, tidy)) %>% unnest(tidy) %>%
  select(name, term:p.value) %>% 
  filter(str_detect(term, "GD") | term=="Geographic_Dist_scaled" | term=="height.cm_scaled") %>%
  drop_na(p.value)

fruits_models_log_CD_GD %>% mutate(rsq=map(lmer, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-f:-glance)
# Can't compute random effect variances. Some variance components equal zero. --> NA conditional rsq
```

## Fruit + Flowers Y2
```{r}
#base models:
FrFlN_modelslog <- tribble(
  ~name,          ~f,
  "1_pop.block",        "logFrFLs ~ (1|pop) + (1|block)", 
  "2_pop.mf.block",     "logFrFLs ~  (1|pop/mf) + (1|block)"
)
#run the models 
FrFlN_modelslog <- FrFlN_modelslog %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_FrFlN_y2_scaled)), #run the models 
         glance = map(lmer, glance)) #glance at the model results
FrFlN_modelslog %>% mutate(tidy=map(lmer, tidy, effects = "ran_pars", scales = "vcov")) %>% unnest(tidy) %>% 
  select(name, group, estimate)

#Add in Geo and tmp/ppt dist
FrFlN_models_log_CD_GD_sub <- tribble(
  ~name,          ~f,
  "1_pop.block",      "logFrFLs ~  height.cm_scaled + (1|pop/mf) + (1|block)", 
  "2_WY_Recent",      "logFrFLs ~  TempDist_Recent_Wtr_Year_2024_scaled + PPTDist_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)",
  "3_WY_Historical",  "logFrFLs ~  TempDist_Historic_Wtr_Year_2024_scaled + PPTDist_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)",
  "4_GS_Recent",      "logFrFLs ~  TempDist_Recent_GrwSsn_2024_scaled + PPTDist_Recent_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)", 
  "5_GS_Historical",  "logFrFLs ~  TempDist_Historic_GrwSsn_2024_scaled + PPTDist_Historic_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)"
)
#run the models 
FrFlN_models_log_CD_GD_sub <- FrFlN_models_log_CD_GD_sub %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_FrFlN_y2_scaled)), #run the models 
         glance = map(lmer, glance)) #glance at the model results
#boundary (singular) fit: see help('isSingular') warning for all models with climate distance 
##show results:
FrFlN_models_log_CD_GD_sub %>% mutate(tidy=map(lmer, tidy)) %>% unnest(tidy) %>%
  select(name, term:p.value) %>% 
  filter(str_detect(term, "Dist") | term=="height.cm_scaled") %>%
  drop_na(p.value)

FrFlN_models_log_CD_GD_sub %>% mutate(rsq=map(lmer, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-f:-glance)
# Can't compute random effect variances. Some variance components equal zero. --> NA conditional rsq

#Add in Geo and Gowers clim dist
FrFlN_models_log_CD_GD <- tribble(
  ~name,          ~f,
  "1_pop.block",      "logFrFLs ~  height.cm_scaled + (1|pop/mf) + (1|block)", 
  "2_WY_Recent",      "logFrFLs ~  GD_Recent_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)",
  "3_WY_Historical",  "logFrFLs ~  GD_Historic_Wtr_Year_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)",
  "4_GS_Recent",      "logFrFLs ~  GD_Recent_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)", 
  "5_GS_Historical",  "logFrFLs ~  GD_Historic_GrwSsn_2024_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block)"
)
#run the models 
FrFlN_models_log_CD_GD <- FrFlN_models_log_CD_GD %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_FrFlN_y2_scaled)), #run the models 
         glance = map(lmer, glance)) #glance at the model results
##show results:
FrFlN_models_log_CD_GD %>% mutate(tidy=map(lmer, tidy)) %>% unnest(tidy) %>%
  select(name, term:p.value) %>% 
  filter(str_detect(term, "GD") | term=="Geographic_Dist_scaled" | term=="height.cm_scaled") %>%
  drop_na(p.value)

FrFlN_models_log_CD_GD %>% mutate(rsq=map(lmer, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-f:-glance)
# Can't compute random effect variances. Some variance components equal zero. --> NA conditional rsq
```

## Prob Fruits Models
```{r}
#basic model workflow:
prob_fitness_wflow <- workflow() %>% 
  add_variables(outcomes = ProbFitness, predictors = c(pop, mf, block))
prob_fitness_fits <- tibble(wflow=list(
  
  pop.block = {prob_fitness_wflow %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ (1|pop) + (1|block))},
  
  pop.mf.block = {prob_fitness_wflow %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
prob_fitness_fits <- prob_fitness_fits %>%
  mutate(fit = map(wflow, fit, data = wl2_prob_fitness_scaled))
prob_fitness_fits %>% pull(fit) 
#convergence issue with pop.block model 

#Add in geo and temp/ppt distance
prob_fitness_SUB_wflow_wl2 <- workflow() %>%
  add_variables(outcomes = ProbFitness, predictors = c(pop, mf, height.cm_scaled,  block, contains("Dist"))) 
prob_fitness_SUB_fits_wl2 <- tibble(wflow=list(
  pop.block = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgTempDist_Recent_Wtr_Year_scaled + AvgPPTDist_Recent_Wtr_Year_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgTempDist_Historic_Wtr_Year_scaled + AvgPPTDist_Historic_Wtr_Year_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  GS_Recent = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgTempDist_Recent_GrwSsn_scaled + AvgPPTDist_Recent_GrwSsn_scaled+ Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  GS_Historical = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgTempDist_Historic_GrwSsn_scaled + AvgPPTDist_Historic_GrwSsn_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent_height = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgTempDist_Recent_Wtr_Year_scaled + AvgPPTDist_Recent_Wtr_Year_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical_height = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgTempDist_Historic_Wtr_Year_scaled + AvgPPTDist_Historic_Wtr_Year_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  GS_Recent_height = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgTempDist_Recent_GrwSsn_scaled + AvgPPTDist_Recent_GrwSsn_scaled+ Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  GS_Historical_height = {prob_fitness_SUB_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgTempDist_Historic_GrwSsn_scaled + AvgPPTDist_Historic_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_prob_fitness_scaled))

##show results:
prob_fitness_SUB_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "Dist") | term=="height.cm_scaled") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)

prob_fitness_SUB_fits_wl2 %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>% 
  mutate(rsq=map(extracted_fit, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-wflow:-extracted_fit)

##predictions for figures:
prob_fitness_sub_preds <- prob_fitness_SUB_fits_wl2 %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="WY_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, 
                         new_data=pred_df_recent_wtr_yr_avgtemp)) %>% 
  unnest(predictions) %>% 
  bind_cols(pred_df_recent_wtr_yr_avgtemp) %>% 
  select(name, .pred, AvgTempDist_Recent_Wtr_Year, AvgPPTDist_Recent_Wtr_Year, Geographic_Dist_scaled) 
#double check that above gives predictions for fixed effects only:
WY_Recent_sub_mod <- glmer(ProbFitness ~ AvgTempDist_Recent_Wtr_Year_scaled +
                             AvgPPTDist_Recent_Wtr_Year_scaled + Geographic_Dist_scaled + 
                             (1|pop/mf) + (1|block),
                           family="binomial", data=wl2_prob_fitness_scaled)
prob_fitness_sub_preds_check <- tibble(.pred=predict(WY_Recent_sub_mod, 
                                                     newdata=pred_df_recent_wtr_yr_avgtemp,
                                               re.form= NA, type="response")) %>% 
  bind_cols(pred_df_recent_wtr_yr_avgtemp) %>% 
  select(.pred, AvgTempDist_Recent_Wtr_Year, AvgPPTDist_Recent_Wtr_Year, Geographic_Dist_scaled)
all.equal(prob_fitness_sub_preds$.pred, prob_fitness_sub_preds_check$.pred)
write_csv(prob_fitness_sub_preds, "../Processed.Data/ProbRep_SubPreds.csv")

#Add in geo and Gowers clim distance
prob_fitness_GD_wflow_wl2 <- workflow() %>%
  add_variables(outcomes = ProbFitness, predictors = c(pop, mf, height.cm_scaled,  block, contains("GD"), Geographic_Dist_scaled_scaled)) 
prob_fitness_GD_fits_wl2 <- tibble(wflow=list(
  pop.block = {prob_fitness_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Recent = {prob_fitness_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgGD_Recent_Wtr_Year_scaled + Geographic_Dist_scaled_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  WY_Historical = {prob_fitness_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgGD_Historic_Wtr_Year_scaled + Geographic_Dist_scaled_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  GS_Recent = {prob_fitness_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgGD_Recent_GrwSsn_scaled + Geographic_Dist_scaled_scaled + height.cm_scaled + (1|pop/mf) + (1|block))},
  
  GS_Historical = {prob_fitness_GD_wflow_wl2 %>% 
      add_model(glmer.model_binomial, formula = ProbFitness ~ AvgGD_Historic_GrwSsn_scaled + Geographic_Dist_scaled_scaled + height.cm_scaled + (1|pop/mf) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow) %>%
  mutate(fit = map(wflow, fit, data = wl2_prob_fitness_scaled))

##show results:
prob_fitness_GD_fits_wl2 %>% mutate(tidy=map(fit, tidy)) %>% unnest(tidy) %>%
  filter(str_detect(term, "GD") | term=="Geographic_Dist_scaled_scaled" | term=="height.cm_scaled") %>%
  drop_na(p.value) %>%
  select(-wflow:-group)

prob_fitness_GD_fits_wl2 %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>% 
  mutate(rsq=map(extracted_fit, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-wflow:-extracted_fit)

##predictions for figures:
prob_fitness_GD_preds <- prob_fitness_GD_fits_wl2 %>% mutate(extracted_fit=map(fit, extract_fit_parsnip)) %>%
  filter(name=="WY_Recent") %>% 
  mutate(predictions=map(extracted_fit, predict, new_data=wl2_prob_fitness_scaled)) %>% 
  unnest(predictions) %>% 
  bind_cols(wl2_prob_fitness_scaled) %>% 
  select(name, .pred, AvgGD_Recent_Wtr_Year) 
write_csv(prob_fitness_GD_preds, "../Processed.Data/ProbRep_GDPreds.csv")

prob_fitness_sub_preds
prob_fitness_sub_preds_check
```

## Total Fruits Models 
```{r}
#base models:
rep_output_modelslog <- tribble(
  ~name,          ~f,
  "1_pop.block",        "logTotalFecundity ~ (1|pop) + (1|block)", 
  "2_pop.mf.block",     "logTotalFecundity ~  (1|pop/mf) + (1|block)"
)
#run the models 
rep_output_modelslog <- rep_output_modelslog %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_rep_output_scaled)), #run the models 
         glance = map(lmer, glance)) #glance at the model results
rep_output_modelslog %>% mutate(tidy=map(lmer, tidy, effects = "ran_pars", scales = "vcov")) %>% unnest(tidy) %>% 
  select(name, group, estimate)

#Add in Geo and temp/ppt dist
rep_output_SUB_models_log_CD_GD <- tribble(
  ~name,          ~f,
  "1_pop.block",      "logTotalFecundity ~  height.cm_scaled + (1|pop) + (1|block)", 
  "2_WY_Recent",      "logTotalFecundity ~  AvgTempDist_Recent_Wtr_Year_scaled + AvgPPTDist_Recent_Wtr_Year_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)",
  "3_WY_Historical",  "logTotalFecundity ~  AvgTempDist_Historic_Wtr_Year_scaled + AvgPPTDist_Historic_Wtr_Year_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)",
  "4_GS_Recent",      "logTotalFecundity ~  AvgTempDist_Recent_GrwSsn_scaled + AvgPPTDist_Recent_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)", 
  "5a_GS_Historical",  "logTotalFecundity ~  AvgTempDist_Historic_GrwSsn_scaled + AvgPPTDist_Historic_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)", 
  #Troubleshoot weird GrwSsn_scaled-AvgTemp_Dist results (positive coefficient, but figure indicates it should be neg...):
  "5b_GS_Historical",  "logTotalFecundity ~  AvgTempDist_Historic_GrwSsn_scaled*AvgPPTDist_Historic_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)", 
  "5c_GS_Historical",  "logTotalFecundity ~  AvgTempDist_Historic_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)", 
  "5d_GS_Historical",  "logTotalFecundity ~  AvgPPTDist_Historic_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)"
)
#run the models 
rep_output_SUB_models_log_CD_GD <- rep_output_SUB_models_log_CD_GD %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_rep_output_scaled)), #run the models 
         glance = map(lmer, glance)) #glance at the model results
#some singularity warnings b/c pop explains 0 var in some models 
rep_output_SUB_models_log_CD_GD %>% mutate(tidy=map(lmer, tidy, effects = "ran_pars", scales = "vcov")) %>% unnest(tidy) %>% 
  select(name, group, estimate)
##show results:
rep_output_SUB_models_log_CD_GD %>% mutate(tidy=map(lmer, tidy)) %>% unnest(tidy) %>%
  select(name, term, estimate:p.value) %>% 
  filter(str_detect(term, "Dist") | term=="height.cm_scaled") %>%
  drop_na(p.value)

rep_output_SUB_models_log_CD_GD %>% mutate(rsq=map(lmer, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-f:-glance)
# Can't compute random effect variances. Some variance components equal zero. --> NA conditional rsq

#Add in Geo and Gowers clim dist
rep_output_models_log_CD_GD <- tribble(
  ~name,          ~f,
  "1_pop.block",      "logTotalFecundity ~  height.cm_scaled + (1|pop) + (1|block)", 
  "2_WY_Recent",      "logTotalFecundity ~  AvgGD_Recent_Wtr_Year_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)",
  "3_WY_Historical",  "logTotalFecundity ~  AvgGD_Historic_Wtr_Year_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)",
  "4_GS_Recent",      "logTotalFecundity ~  AvgGD_Recent_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)", 
  "5_GS_Historical",  "logTotalFecundity ~  AvgGD_Historic_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)"
)
#run the models 
rep_output_models_log_CD_GD <- rep_output_models_log_CD_GD %>%
  mutate(lmer = map(f, ~ lmer(as.formula(.), 
                            data = wl2_rep_output_scaled)), #run the models 
         glance = map(lmer, glance)) #glance at the model results
##show results:
rep_output_models_log_CD_GD %>% mutate(tidy=map(lmer, tidy)) %>% unnest(tidy) %>%
  select(name, term:p.value) %>% 
  filter(str_detect(term, "GD") | term=="Geographic_Dist_scaled" | term=="height.cm_scaled") %>%
  drop_na(p.value)

rep_output_models_log_CD_GD %>% mutate(rsq=map(lmer, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-f:-glance)
```
