---
title: "Climate_corrs_pcas"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code for Figures 1 and S1; Table S2 and S3 

To Do:

-   Save a full PCA with loadings for the hulls PCA (will be a third panel in FigS2)

## Load libraries
```{r}
library(tidyverse)
library(ggrepel) #for repelling labels 
library(corrplot) #plotting correlations 
library(vegan) #for permanova 
library(ggfortify) #plotting PCAs
library(viridis) #for color scale

#making topo maps:
library(elevatr)
library(terra)
library(sf)
library(giscoR)
library(marmap)
```

## Load Climate Data
```{r}
all_clim <- read_csv("../Processed.Data/All_Clim.csv")
```

## Map of populations (Figure 1a)
```{r}
pops_garden_locs <- all_clim %>% 
  mutate_at(c("Lat", "Long"), as.double) %>% 
  select(parent.pop:Long) %>%
  filter(!is.na(Lat), !is.na(Long)) %>% 
  distinct()
states <- map_data("state") %>% filter(region == "california")

ggplot() +
  geom_polygon(data = states, aes(x = long, y = lat, group = group), fill = "gray") +
  coord_quickmap(xlim = c(-125, -114), ylim = c(35.8, 41))+
  geom_point(data = pops_garden_locs,
             aes(x = Long, y = Lat, color=elev_m),
             size = 3) +
  geom_label_repel(data = pops_garden_locs,
         aes(x = Long, y = Lat,
             label = `parent.pop`),
         min.segment.length = 0,
         max.overlaps = 100,
         #force = 3,
         box.padding = 0.4,
         label.padding = 0.15,
         label.size = 0.1,
         size = 3) +
  labs(color="Elevation (m)") +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_void()
#ggsave("../Figures/Pop_Map_Figure1a.png", width = 10, height = 10, units = "in")
```

## Climate variable correlations (Figure S1)

### Calculate 30 year climate averages 
```{r}
all_clim_avgs <- all_clim %>% 
  select(-Lat, -Long) %>% 
  filter(parent.pop!="WL2_Garden") %>% #remove the garden 
  left_join(pops_garden_locs) %>% #add in Lat Long for growth season rows 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, 
           timeframe, Season) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx",
                 "ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_month", "tmean_driest_month", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"), 
               c(mean), na.rm = TRUE)
```

### Recent Corrs - Water Year
```{r}
recent_wtr_yr_avgs_normalized <- all_clim_avgs %>% 
  filter(timeframe=="recent", Season=="Water Year") %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_recent_wtr_yr = cor(recent_wtr_yr_avgs_normalized) #test correlations among the traits
cor.sig_recent_wtr_yr <- cor.mtest(recent_wtr_yr_avgs_normalized, method = "pearson") #get pearson's test p-values 

corrplot(cor.norm_recent_wtr_yr, type="upper",
         tl.srt = 45, p.mat = cor.sig_recent_wtr_yr$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

### Historical Corrs - Water year
```{r}
historic_wtr_yr_avgs_normalized <- all_clim_avgs %>% 
  filter(timeframe=="historic", Season=="Water Year") %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_historic_wtr_yr = cor(historic_wtr_yr_avgs_normalized) #test correlations among the traits
cor.sig_historic_wtr_yr <- cor.mtest(historic_wtr_yr_avgs_normalized, method = "pearson") #get pearson's test p-values 

corrplot(cor.norm_historic_wtr_yr, type="upper",
         tl.srt = 45, p.mat = cor.sig_historic_wtr_yr$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

### Recent Corrs - Growth Season
```{r}
recent_grwssn_avgs_normalized <- all_clim_avgs %>% 
  filter(timeframe=="recent", Season=="Growth Season") %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_recent_grwssn = cor(recent_grwssn_avgs_normalized) #test correlations among the traits
cor.sig_recent_grwssn <- cor.mtest(recent_grwssn_avgs_normalized, method = "pearson") #get pearson's test p-values 

corrplot(cor.norm_recent_grwssn, type="upper",
         tl.srt = 45, p.mat = cor.sig_recent_grwssn$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

### Historical Corrs - Growth Season
```{r}
historic_grwssn_avgs_normalized <- all_clim_avgs %>% 
  filter(timeframe=="historic", Season=="Growth Season") %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_historic_grwssn = cor(historic_grwssn_avgs_normalized) #test correlations among the traits
cor.sig_historic_grwssn <- cor.mtest(historic_grwssn_avgs_normalized, method = "pearson") #get pearson's test p-values 

corrplot(cor.norm_historic_grwssn, type="upper",
         tl.srt = 45, p.mat = cor.sig_historic_grwssn$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

## Climate Change PCAs (Figure 1b-c; Table S2, S3)

### Water Year + Growth Season PCA

```{r}
all_clim_avgs_normalized <- all_clim_avgs %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_all_clim = cor(all_clim_avgs_normalized) #test correlations among the traits
cor.sig_all_clim <- cor.mtest(all_clim_avgs_normalized, method = "pearson") #get pearson's test p-values 
cor.norm_all_clim
cor.sig_all_clim$p
#tmn, tmx, and ann_tmean all highly correlated (~99%) - only keep ann_tmean 
#tmean_wettest_month, tmean_driest_month correlated to ann_tmean (92-93%) #low enough that it's okay to keep in?
all_clim_avgs <- all_clim_avgs %>% 
  ungroup()
all_clim_avgs.pc = prcomp(all_clim_avgs[c(8:10, 13:22)], scale = TRUE, center = TRUE)
summary(all_clim_avgs.pc)
tibble(PC=str_c("PC",str_pad(1:13,2,pad="0")),
       percent_var=all_clim_avgs.pc$sdev[1:13]^2/sum(all_clim_avgs.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")

#combine pcs with metadata
all_clim_avgs.pc.dat = data.frame(all_clim_avgs.pc$x)
all_clim_avgs_locs.pc = cbind(all_clim_avgs, all_clim_avgs.pc.dat) %>% 
  select(parent.pop, elev_m, , Lat, timeframe, Season, PC1:PC13)
all_clim_avgs_loadings = data.frame(varnames=rownames(all_clim_avgs.pc$rotation), all_clim_avgs.pc$rotation)
all_clim_avgs_loadings

#PC1 pos ann_tmean loading
#PC2 pos for ann_ppt
```

#### Permanova 
```{r}
all_clim_avgs_locs.pc_dist <- all_clim_avgs_locs.pc %>% ungroup() %>% select(PC1:PC13)
dist_matrix_allclim <- dist(all_clim_avgs_locs.pc_dist, method = "euclidian") #use a distance function to calculate euclidian distance in PCA space
permanova_results_allclim <- adonis2(dist_matrix_allclim ~ Season*timeframe*elev_m*Lat, data = all_clim_avgs_locs.pc) #use adonis2 to run the permanova
permanova_results_allclim #look at output 
#get stats per term in the model:
permanova_results_allclim_terms <- adonis2(dist_matrix_allclim ~ Season*timeframe*elev_m*Lat, data = all_clim_avgs_locs.pc, by = "terms")
permanova_results_allclim_terms

#LM on PCs (follow up on permanova)
lmer_results_allclim <- all_clim_avgs_locs.pc %>%
  ungroup() %>% 
  pivot_longer(starts_with("PC", ignore.case = FALSE), 
               names_to = "PC", values_to = "value") %>% 
  group_by(PC) %>% 
  nest(data=c(Season, timeframe, parent.pop, elev_m, Lat, value)) %>% 
  mutate(glm=map(data, ~ glm(value ~ Season*timeframe*elev_m*Lat,
                               data=.x)),
         anova = map(glm, ~ broom.mixed::tidy(anova(.x))))

PC_anova_all_clim <- lmer_results_allclim %>% select(-data, -glm) %>% unnest(anova) %>%
  select(PC, term, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(PC, p.value)
PC_anova_all_clim #PC1 and PC2 have significant timeframe effects 
#PC1 has sig timeframe*elev_m effect too

lmer_results_allclim %>% select(-data, -glm) %>% unnest(anova) %>% filter(PC=="PC1" | PC=="PC2" |PC=="PC3" | PC=="PC4") 
```

#### Plot climate change PCA

```{r}
autoplot(all_clim_avgs.pc, data = all_clim_avgs,
         x=1, y=2,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
```

```{r}
home_sites_pca_allclim <- all_clim_avgs_locs.pc %>%  
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=timeframe, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  labs(x="PC1 (36.47%)", y="PC2 (25.94%)", color="Elevation (m)") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .9) +
  theme_classic() +
  theme(text=element_text(size=28)) +
  coord_cartesian(clip = "off") +
  facet_wrap(~Season)

## add WL2 garden 2023 and 2024
WL2GRDN_allclim_pc_prep_2023_grwssn <- all_clim %>% 
  ungroup() %>% 
  filter(parent.pop=="WL2_Garden", year==2023, Season=="Growth Season") %>% 
  select(cwd:ppt, ann_tmean:ppt_coldest_month) 
WL2GRDN_allclim_predicted_2023_grwssn <- predict(all_clim_avgs.pc, newdata = WL2GRDN_allclim_pc_prep_2023_grwssn)
WL2GRDN_allclim_pc_prep_2023_wtryr <- all_clim %>% 
  ungroup() %>% 
  filter(parent.pop=="WL2_Garden", year==2023, Season=="Water Year") %>% 
  select(cwd:ppt, ann_tmean:ppt_coldest_month) 
WL2GRDN_allclim_predicted_2023_wtryr <- predict(all_clim_avgs.pc, newdata = WL2GRDN_allclim_pc_prep_2023_wtryr)

WL2GRDN_allclim_pc_prep_2024_grwssn <- all_clim %>% 
  ungroup() %>% 
  filter(parent.pop=="WL2_Garden", year==2024, Season=="Growth Season") %>% 
  select(cwd:ppt, ann_tmean:ppt_coldest_month) 
WL2GRDN_allclim_predicted_2024_grwssn <- predict(all_clim_avgs.pc, newdata = WL2GRDN_allclim_pc_prep_2024_grwssn)
WL2GRDN_allclim_pc_prep_2024_wtryr <- all_clim %>% 
  ungroup() %>% 
  filter(parent.pop=="WL2_Garden", year==2024, Season=="Water Year") %>% 
  select(cwd:ppt, ann_tmean:ppt_coldest_month) 
WL2GRDN_allclim_predicted_2024_wtryr <- predict(all_clim_avgs.pc, newdata = WL2GRDN_allclim_pc_prep_2024_wtryr)

home_sites_pca_allclim$data <- rbind(home_sites_pca_allclim$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elev_m = 2020,
    Lat = NA,
    timeframe = c("2023", "2023", "2024", "2024"),
    Season = c("Growth Season", "Water Year", "Growth Season", "Water Year"),
    PC1 =  c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC1"],  WL2GRDN_allclim_predicted_2023_wtryr[, "PC1"] , WL2GRDN_allclim_predicted_2024_grwssn[, "PC1"],  WL2GRDN_allclim_predicted_2024_wtryr[, "PC1"] ),
    PC2 =  c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC2"],  WL2GRDN_allclim_predicted_2023_wtryr[, "PC2"] , WL2GRDN_allclim_predicted_2024_grwssn[, "PC2"],  WL2GRDN_allclim_predicted_2024_wtryr[, "PC2"] ),
    PC3 =  c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC3"],  WL2GRDN_allclim_predicted_2023_wtryr[, "PC3"] , WL2GRDN_allclim_predicted_2024_grwssn[, "PC3"],  WL2GRDN_allclim_predicted_2024_wtryr[, "PC3"] ),
    PC4 =  c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC4"],  WL2GRDN_allclim_predicted_2023_wtryr[, "PC4"] , WL2GRDN_allclim_predicted_2024_grwssn[, "PC4"],  WL2GRDN_allclim_predicted_2024_wtryr[, "PC4"] ),
    PC5 =  c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC5"],  WL2GRDN_allclim_predicted_2023_wtryr[, "PC5"] , WL2GRDN_allclim_predicted_2024_grwssn[, "PC5"],  WL2GRDN_allclim_predicted_2024_wtryr[, "PC5"] ),
    PC6 =  c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC6"],  WL2GRDN_allclim_predicted_2023_wtryr[, "PC6"] , WL2GRDN_allclim_predicted_2024_grwssn[, "PC6"],  WL2GRDN_allclim_predicted_2024_wtryr[, "PC6"] ),
    PC7 =  c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC7"],  WL2GRDN_allclim_predicted_2023_wtryr[, "PC7"] , WL2GRDN_allclim_predicted_2024_grwssn[, "PC7"],  WL2GRDN_allclim_predicted_2024_wtryr[, "PC7"] ),
    PC8 =  c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC8"],  WL2GRDN_allclim_predicted_2023_wtryr[, "PC8"] , WL2GRDN_allclim_predicted_2024_grwssn[, "PC8"],  WL2GRDN_allclim_predicted_2024_wtryr[, "PC8"] ),
    PC9 =  c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC9"],  WL2GRDN_allclim_predicted_2023_wtryr[, "PC9"] , WL2GRDN_allclim_predicted_2024_grwssn[, "PC9"],  WL2GRDN_allclim_predicted_2024_wtryr[, "PC9"] ),
    PC10 = c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC10"], WL2GRDN_allclim_predicted_2023_wtryr[, "PC10"], WL2GRDN_allclim_predicted_2024_grwssn[, "PC10"], WL2GRDN_allclim_predicted_2024_wtryr[, "PC10"]),
    PC11 = c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC11"], WL2GRDN_allclim_predicted_2023_wtryr[, "PC11"], WL2GRDN_allclim_predicted_2024_grwssn[, "PC11"], WL2GRDN_allclim_predicted_2024_wtryr[, "PC11"]),
    PC12 = c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC12"], WL2GRDN_allclim_predicted_2023_wtryr[, "PC12"], WL2GRDN_allclim_predicted_2024_grwssn[, "PC12"], WL2GRDN_allclim_predicted_2024_wtryr[, "PC12"]),
    PC13 = c(WL2GRDN_allclim_predicted_2023_grwssn[, "PC13"], WL2GRDN_allclim_predicted_2023_wtryr[, "PC13"], WL2GRDN_allclim_predicted_2024_grwssn[, "PC13"], WL2GRDN_allclim_predicted_2024_wtryr[, "PC13"]),
    group = c("new", "new2", "new3", "new4")
  )
)

WL2_label <- data.frame(
  label = c("WL2 \n 2023", "WL2 \n 2024",
            "WL2 \n 2023", "WL2 \n 2024"),
  Season   = c("Growth Season", "Growth Season",
               "Water Year", "Water Year"),
  x     = c(-0.9, 2.5,
            -6.0, -1.5),
  y     = c(-0.92, -3.5,
            5.6, 3),
  timeframe = NA
)

home_sites_pca_allclim + 
  geom_point(data=filter(home_sites_pca_allclim$data, parent.pop == "WL2_Garden"), 
             size=6, shape = 18, show.legend = FALSE) +
  geom_text(data = WL2_label,
  mapping = aes(x = x, y = y, label = label),
  colour = "#8959c9",
  size=7
) +
  geom_path(data=filter(home_sites_pca_allclim$data, parent.pop == "WL2"),
            aes(group=group),arrow = arrow(length=unit(5, "points"), 
                                           type="closed"), linewidth = 1.9) 
ggsave("../Figures/AllClim_PC1-PC2_PlusGarden.png", width = 14.8, height = 6, units = "in")
```

## All climate PCA (all years, water year, growing season, historic and recent) - Fig S2

```{r}
all_clim = all_clim %>%
  #remove data before historic (which should be just WL2_Garden)
  filter(year >= 1964) %>%
  #create time frames for WL2 garden site
  mutate(timeframe = if_else(parent.pop == "WL2_Garden" & year < 1994, "Historic", timeframe)) %>%
  mutate(timeframe = if_else(parent.pop == "WL2_Garden" & year >= 1994 & year != 2024, "recent", timeframe)) %>%
  mutate(timeframe = if_else(parent.pop == "WL2_Garden" & year == 2024, "2024Garden", timeframe))%>%
  mutate(timeframe = if_else(parent.pop == "WL2_Garden" & year == 2023, "2023Garden", timeframe)) %>%
  mutate(timeframe = as.factor(timeframe), Season = as.factor(Season), 
         parent.pop = as.factor(parent.pop), elevation.group = as.factor(elevation.group)) %>%
    #combine season and timeframe 
  mutate(seasontime = as.factor(paste(Season, timeframe, sep = "_"))) %>% 
  mutate(timeframe=if_else(timeframe=="recent", "Recent",
                 if_else(timeframe=="historic", "Historic", timeframe)))
#summary(all_clim)

allclim_normalized <- all_clim %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_all = cor(allclim_normalized) #test correlations among the traits
cor.sig_all <- cor.mtest(allclim_normalized, method = "pearson") #get pearson's test p-values 
cor.norm_all
cor.sig_all$p
#tmn and tmx highly correlated with ann_tmean (~98%)   - only keep ann_tmean 
#JRG: keep all for comparisons between water year, etc? 

allclim.pc = prcomp(all_clim[c(8:10, 13:22)], scale = TRUE, center = TRUE) #removed tmn and tmx 
summary(allclim.pc)
tibble(PC=str_c("PC",str_pad(1:13,2,pad="0")),
       percent_var=allclim.pc$sdev[1:13]^2/sum(allclim.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")

#combine pcs with metadata
allclim.pc.dat = data.frame(allclim.pc$x)
allclim_locs.pc = cbind(all_clim, allclim.pc.dat)
allclim_loadings = data.frame(varnames=rownames(allclim.pc$rotation), allclim.pc$rotation) %>% mutate(elev_m=0)
allclim_loadings

#PC1 pos snowpack, pos ppt and ann_ppt, and negative ann_tmean
#PC2 pos tmean_driest_month, pos temp_ann_range, pos ann_ppt 
```

### Plot PCA with all climate

```{r}
##plot everything, color by elevation
autoplot(allclim.pc, data = all_clim,
         x=1, y=2,
         colour='elev_m', alpha=0.5,
         label=FALSE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()
ggsave("../Figures/AllClim_YrlyVar.png", width = 7.4, height = 6, units = "in")

#facet by seasonal summmary 
autoplot(allclim.pc, data = all_clim,
         x=1, y=2,
         colour='elev_m', alpha=0.5,
         label=FALSE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic() +
  facet_wrap(~Season)

#facet by timeperiod 
autoplot(allclim.pc, data = all_clim,
         x=1, y=2,
         colour='elev_m', alpha=0.5,
         label=FALSE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic() +
  facet_wrap(~timeframe)
#this one isn't pretty because of the garden 
```

### Hulls prep

```{r}
allclim_locs.pc_nogarden = allclim_locs.pc %>%
                          filter(parent.pop != "WL2_Garden")
allclim_locs.pc_garden = allclim_locs.pc %>%
  filter(parent.pop == "WL2_Garden", year>2022) %>% 
  mutate(year=as.character(year))

#plot with no vectors 
pca_all_novect12 =  ggplot(allclim_locs.pc_nogarden, aes(x = PC1, y = PC2, colour= elev_m)) +
  geom_point(aes(colour = elev_m), size = 3, alpha=0.5) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  labs(x="PC1 (29.74%)", y="PC2 (20.11%)", color="Elevation (m)") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  annotate("text", x = -3, y = -6.4, label = "Warm/No Snow", size=5) +
  annotate("text", x = 7.4, y = -6.4, label = "Cold/Snow", size=5) +
  theme_classic() +
  theme(text=element_text(size=28))
pca_all_novect12
```

### Hulls by water year vs growth season
```{r}
pca_hull_season_12 <- 
  allclim_locs.pc_nogarden %>% 
  group_by(Season, timeframe) %>% 
  slice(chull(PC1, PC2))

pca_chull_season = pca_all_novect12 +
  geom_polygon(data = pca_hull_season_12,
               aes(x = PC1, y = PC2, group = Season, linetype = Season, fill = Season), 
               color = "black", alpha = 0.3, show.legend = T) + 
  scale_fill_manual(values = c("gray72", "gray22")) +
  facet_wrap(~timeframe)
pca_chull_season

#add back in WL2 garden values
allclim_locs.pc_garden_2 <- allclim_locs.pc_garden %>% 
  mutate(timeframe="Recent")
allclim_locs.pc_garden_3 <- allclim_locs.pc_garden %>% 
  mutate(timeframe="Historic")
allclim_locs.pc_garden_4 <- bind_rows(allclim_locs.pc_garden_2, allclim_locs.pc_garden_3) %>% 
  mutate(seasontime = str_remove(seasontime, "Garden"))

#labels for y axis
Y_label <- data.frame(
  label = c("Low Temp Range", "High Temp Range",
            "High PPT in", "Low PPT in",
            "Warmest Month", "Warmest Month"),
  timeframe = c("Historic", "Historic",
             "Historic", "Historic",
             "Historic", "Historic"),
  x     = c(-3.3, -3.3,
            -4.1, -4.1,
            -3.5, -3.5),
  y     = c(-3.7, 4.5,
            -4.7, 3.5,
            -5.2, 3)
)

pca_chull_season_wgarden = pca_chull_season + 
  geom_point(data = allclim_locs.pc_garden_4, aes(shape=seasontime),
              size=6, color = "purple")  +
  scale_shape_manual(values=c(15, 16, 17, 18), name = "WL2 Garden") +
  geom_text(data = Y_label,
  mapping = aes(x = x, y = y, label = label),
  colour = "black",
  size=5
)

#for adding all loading arrows:
  #geom_segment(data = allclim_loadings, aes(x=0, y=0, xend = PC1*5, yend = PC2*5), #0.8 just scales loading arrows to fit graph, autoplot does this automatically when plotting loading = T
               #      arrow = arrow(length = unit(0.2, "cm"), type= "closed"),  size = 0.75, color = "black")  +
   #     geom_text(data = allclim_loadings, mapping = aes(label = varnames, x = PC1*5.4, y=PC2*5.4), color = "black")  #multiply by 5.4 to move a little away from arrow
pca_chull_season_wgarden
ggsave("../Figures/AllClim_YrlyVar_SeasonHulls.png", width = 14.8, height = 6, units = "in")
```