---
title: "Climate_corrs_pcas"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code for Figures 1 and S1; Table S2 and S3 

## Load libraries
```{r}
library(tidyverse)
library(ggrepel) #for repelling labels 
library(corrplot) #plotting correlations 
library(vegan) #for permanova 
library(ggfortify) #plotting PCAs

#making topo maps:
library(elevatr)
library(terra)
library(sf)
library(giscoR)
library(marmap)
```

## Load Climate Data
```{r}
all_clim <- read_csv("../Processed.Data/Climate/All_Clim.csv")
```

## Map of populations (Figure 1a)
```{r}
pops_garden_locs <- all_clim %>% 
  mutate_at(c("Lat", "Long"), as.double) %>% 
  select(parent.pop:Long) %>%
  filter(!is.na(Lat), !is.na(Long)) %>% 
  distinct()
states <- map_data("state") %>% filter(region == "california")

ggplot() +
  geom_polygon(data = states, aes(x = long, y = lat, group = group), fill = "gray") +
  coord_quickmap(xlim = c(-125, -114), ylim = c(35.8, 41))+
  geom_point(data = pops_garden_locs,
             aes(x = Long, y = Lat, color=elev_m),
             size = 3) +
  geom_label_repel(data = pops_garden_locs,
         aes(x = Long, y = Lat,
             label = `parent.pop`),
         min.segment.length = 0,
         max.overlaps = 100,
         #force = 3,
         box.padding = 0.4,
         label.padding = 0.15,
         label.size = 0.1,
         size = 3) +
  labs(color="Elevation (m)") +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  theme_void()
ggsave("../Figures/Pop_Map_Figure1a.png", width = 10, height = 10, units = "in")
```

## Climate variable correlations (Figure S1)

### Calculate 30 year climate averages 
```{r}
all_clim_avgs <- all_clim %>% 
  select(-Lat, -Long) %>% 
  filter(parent.pop!="WL2_Garden") %>% #remove the garden 
  left_join(pops_garden_locs) %>% #add in Lat Long for growth season rows 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, 
           timeframe, Season) %>% 
  summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx",
                 "ann_tmean", "mean_diurnal_range", "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_month", "tmean_driest_month", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"), 
               c(mean), na.rm = TRUE)
```

### Recent Corrs - Water Year
```{r}
recent_wtr_yr_avgs_normalized <- all_clim_avgs %>% 
  filter(timeframe=="recent", Season=="Water Year") %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_recent_wtr_yr = cor(recent_wtr_yr_avgs_normalized) #test correlations among the traits
cor.sig_recent_wtr_yr <- cor.mtest(recent_wtr_yr_avgs_normalized, method = "pearson") #get pearson's test p-values 

corrplot(cor.norm_recent_wtr_yr, type="upper",
         tl.srt = 45, p.mat = cor.sig_recent_wtr_yr$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

### Historical Corrs - Water year
```{r}
historic_wtr_yr_avgs_normalized <- all_clim_avgs %>% 
  filter(timeframe=="historic", Season=="Water Year") %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_historic_wtr_yr = cor(historic_wtr_yr_avgs_normalized) #test correlations among the traits
cor.sig_historic_wtr_yr <- cor.mtest(historic_wtr_yr_avgs_normalized, method = "pearson") #get pearson's test p-values 

corrplot(cor.norm_historic_wtr_yr, type="upper",
         tl.srt = 45, p.mat = cor.sig_historic_wtr_yr$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

### Recent Corrs - Growth Season
```{r}
recent_grwssn_avgs_normalized <- all_clim_avgs %>% 
  filter(timeframe=="recent", Season=="Growth Season") %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_recent_grwssn = cor(recent_grwssn_avgs_normalized) #test correlations among the traits
cor.sig_recent_grwssn <- cor.mtest(recent_grwssn_avgs_normalized, method = "pearson") #get pearson's test p-values 

corrplot(cor.norm_recent_grwssn, type="upper",
         tl.srt = 45, p.mat = cor.sig_recent_grwssn$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

### Historical Corrs - Growth Season
```{r}
historic_grwssn_avgs_normalized <- all_clim_avgs %>% 
  filter(timeframe=="historic", Season=="Growth Season") %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_historic_grwssn = cor(historic_grwssn_avgs_normalized) #test correlations among the traits
cor.sig_historic_grwssn <- cor.mtest(historic_grwssn_avgs_normalized, method = "pearson") #get pearson's test p-values 

corrplot(cor.norm_historic_grwssn, type="upper",
         tl.srt = 45, p.mat = cor.sig_historic_grwssn$p, 
         sig.level = 0.05, insig="blank")
#800 x 734
```

## Climate Change PCAs (Figure 1b-c; Table S2, S3)

### Water Year PCA
```{r}
wtr_yr_avgs_normalized <- all_clim_avgs %>% 
  filter(Season=="Water Year") %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_wtr_yr = cor(wtr_yr_avgs_normalized) #test correlations among the traits
cor.sig_wtr_yr <- cor.mtest(wtr_yr_avgs_normalized, method = "pearson") #get pearson's test p-values 
cor.norm_wtr_yr
cor.sig_wtr_yr$p
#ann_ppt and ppt 100% correlated (ppt = avg across monts, ann_ppt = avg of the total ppt in a year) - only keep ann_ppt 
#tmn, tmx, tmean_wettest_month, tmean_driest_month and ann_tmean all highly correlated (97-99%) - only keep ann_tmean 
#ppt warmest month highly neg correlated with tmn, ann_tmean, tmean_driest - take it out 
#temp ann range and mean diurnal range highly corr - keep temp_ann_range

wtr_yr_avgs <- all_clim_avgs %>% 
  filter(Season=="Water Year") %>% 
  ungroup()
wtr_yr_avgs.pc = prcomp(wtr_yr_avgs[c(8:9, 13, 15:16, 19:20, 22)], scale = TRUE, center = TRUE)
summary(wtr_yr_avgs.pc)
tibble(PC=str_c("PC",str_pad(1:8,2,pad="0")),
       percent_var=wtr_yr_avgs.pc$sdev[1:8]^2/sum(wtr_yr_avgs.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")

#combine pcs with metadata
wtr_yr_avgs.pc.dat = data.frame(wtr_yr_avgs.pc$x)
wtr_yr_avgs_locs.pc = cbind(wtr_yr_avgs, wtr_yr_avgs.pc.dat)
wtr_yr_avgs_loadings = data.frame(varnames=rownames(wtr_yr_avgs.pc$rotation), wtr_yr_avgs.pc$rotation)
wtr_yr_avgs_loadings
```

#### Permanova 
```{r}
wtr_yr_avgs_locs.pc_dist <- wtr_yr_avgs_locs.pc %>% ungroup() %>% select(PC1:PC8)
dist_matrix_wtr_year <- dist(wtr_yr_avgs_locs.pc_dist, method = "euclidian") #use a distance function to calculate euclidian distance in PCA space
permanova_results_wtr_year <- adonis2(dist_matrix_wtr_year ~ timeframe*elev_m*Lat, data = wtr_yr_avgs_locs.pc) #use adonis2 to run the permanova
permanova_results_wtr_year #look at output 
#get stats per term in the model:
permanova_results_wtr_year_terms <- adonis2(dist_matrix_wtr_year ~ timeframe*elev_m*Lat, data = wtr_yr_avgs_locs.pc, by = "terms")
permanova_results_wtr_year_terms

#LM on PCs (follow up on permanova)
lmer_results_wtr_year <- wtr_yr_avgs_locs.pc %>%
  ungroup() %>% 
  select(timeframe, parent.pop, elev_m, Lat, Long, PC1:PC8) %>% 
  pivot_longer(starts_with("PC", ignore.case = FALSE), 
               names_to = "PC", values_to = "value") %>% 
  group_by(PC) %>% 
  nest(data=c(timeframe, parent.pop, elev_m, Lat, Long, value)) %>% 
  mutate(glm=map(data, ~ glm(value ~ timeframe*elev_m*Lat,
                               data=.x)),
         anova = map(glm, ~ broom.mixed::tidy(anova(.x))))

PC_anova_wtr_yr <- lmer_results_wtr_year %>% select(-data, -glm) %>% unnest(anova) %>%
  select(PC, term, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(term, p.value)
PC_anova_wtr_yr #PC2 and PC4 most sig, PC1 and PC7 also have sig timeframe effects 

lmer_results_wtr_year %>% select(-data, -glm) %>% unnest(anova) %>% filter(PC=="PC1" | PC=="PC2" |PC=="PC3" | PC=="PC4") 
```

#### Plot climate change PCA
```{r}
#prep the pc for making a plot with arrows distinguishing recent vs. historical time 
wtr_yr_avgs_locs.pc_avg <- wtr_yr_avgs_locs.pc %>%
  group_by(parent.pop, elev_m, timeframe) %>%
  summarise(across(.cols=starts_with("PC", ignore.case = FALSE), .fns = mean)) %>%
  ungroup()

autoplot(wtr_yr_avgs.pc, data = wtr_yr_avgs,
         x=1, y=2,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

autoplot(wtr_yr_avgs.pc, data = wtr_yr_avgs,
         x=2, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

autoplot(wtr_yr_avgs.pc, data = wtr_yr_avgs,
         x=1, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

wtr_yr_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=timeframe, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

wtr_yr_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC2, y=PC4, shape=timeframe, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

wtr_yr_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC4, shape=timeframe, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

#wtr_yr_avgs_locs.pc_avg %>% filter(parent.pop=="WL2")
#wtr_yr_avgs_loadings
```

```{r}
#PC1 = positive PPT and PCK, negative ANN_TMEAN
#PC2 = positive ppt_seasonality and ppt_coldest month, negative temp_ann_range
#PC4 = positive CWD and temp_seasonality and negative ann_tmean
home_sites_pca_wtryr <- wtr_yr_avgs_locs.pc_avg %>%  
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=timeframe, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  labs(x="PC1 (50.25%)", y="PC2 (18.25%)", color="Elevation (m)") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  annotate("text", x = 1.3, y= 1.16, label = "WL2", colour = "purple", size=7) +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8) +
  annotate("text", x = -4.5, y = -5.4, label = "Warm \n No Snow", size=6) +
  annotate("text", x = 4.5, y = -5.4, label = "Cold \n Snow", size=6) +
  annotate("text", x = -6.2, y = -3, label = "Low PPT \n Seasonality", size=6) +
  annotate("text", x = -6.2, y = 3, label = "High PPT \n Seasonality", size=6) +
  coord_cartesian(ylim = c(-4, 4), xlim = c(-5,5), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))

## add WL2 garden 2023 and 2024
WL2GRDN_wtryr_pc_prep_2023 <- all_clim %>% 
  ungroup() %>% 
  filter(parent.pop=="WL2_Garden", year==2023, Season=="Water Year") %>% 
  select(cwd:pck, ann_tmean, temp_seasonality, temp_ann_range, ann_ppt, ppt_seasonality, ppt_coldest_month) 
WL2GRDN_wtryr_predicted_2023 <- predict(wtr_yr_avgs.pc, newdata = WL2GRDN_wtryr_pc_prep_2023)

WL2GRDN_wtryr_pc_prep_2024 <- all_clim %>% 
  ungroup() %>% 
  filter(parent.pop=="WL2_Garden", year==2024, Season=="Water Year") %>% #need 2024 water year data for WL2 garden 
  select(cwd:pck, ann_tmean, temp_seasonality, temp_ann_range, ann_ppt, ppt_seasonality, ppt_coldest_month) 
WL2GRDN_wtryr_2024_predicted <- predict(wtr_yr_avgs.pc, newdata = WL2GRDN_wtryr_pc_prep_2024)

home_sites_pca_wtryr$data <- rbind(home_sites_pca_wtryr$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elev_m = 2020,
    timeframe = c("2023", "2024"),
    PC1 = c(WL2GRDN_wtryr_predicted_2023[, "PC1"], WL2GRDN_wtryr_2024_predicted[, "PC1"]),
    PC2 = c(WL2GRDN_wtryr_predicted_2023[, "PC2"], WL2GRDN_wtryr_2024_predicted[, "PC2"]),
    PC3 = c(WL2GRDN_wtryr_predicted_2023[, "PC3"], WL2GRDN_wtryr_2024_predicted[, "PC3"]),
    PC4 = c(WL2GRDN_wtryr_predicted_2023[, "PC4"], WL2GRDN_wtryr_2024_predicted[, "PC4"]),
    PC5 = c(WL2GRDN_wtryr_predicted_2023[, "PC5"], WL2GRDN_wtryr_2024_predicted[, "PC5"]),
    PC6 = c(WL2GRDN_wtryr_predicted_2023[, "PC6"], WL2GRDN_wtryr_2024_predicted[, "PC6"]),
    PC7 = c(WL2GRDN_wtryr_predicted_2023[, "PC7"], WL2GRDN_wtryr_2024_predicted[, "PC7"]),
    PC8 = c(WL2GRDN_wtryr_predicted_2023[, "PC8"], WL2GRDN_wtryr_2024_predicted[, "PC8"]),
    group = c("new", "new2")
  )
)

home_sites_pca_wtryr + 
  geom_point(data=filter(home_sites_pca_wtryr$data, parent.pop == "WL2_Garden"), size=6, shape = 18, show.legend = FALSE) +
  annotate("text", x = 4.49, y= 2.7, label = "WL2 \n Garden \n 2023", colour = "purple", size=6) +
  annotate("text", x = -0.71, y= 1.9, label = "WL2 \n Garden \n 2024", colour = "purple", size=6) 
ggsave("../Figures/Wtr_Year_PC1-PC2_PlusGarden.png", width = 7.4, height = 6, units = "in")
```


### Growth Season PCA
```{r}
grwssn_avgs_normalized <- all_clim_avgs %>% 
  filter(Season=="Growth Season") %>% 
  ungroup() %>% 
  select(cwd:ppt_coldest_month) %>% 
  scale() #center and scale the data 

cor.norm_grwssn = cor(grwssn_avgs_normalized) #test correlations among the traits
cor.sig_grwssn <- cor.mtest(grwssn_avgs_normalized, method = "pearson") #get pearson's test p-values 
cor.norm_grwssn
cor.sig_grwssn$p
#tmn, tmx, tmean_driest_month and ann_tmean all highly correlated (90-98%) - only keep ann_tmean 
#ppt_warmest_month highly neg corr with tmn, ann_tmean - take it out 

grwssn_avgs <- all_clim_avgs %>% 
  filter(Season=="Growth Season") %>% 
  ungroup()
grwssn_avgs.pc = prcomp(grwssn_avgs[c(8:10, 13:17, 19:20, 22)], scale = TRUE, center = TRUE)
summary(grwssn_avgs.pc)
tibble(PC=str_c("PC",str_pad(1:11,2,pad="0")),
       percent_var=grwssn_avgs.pc$sdev[1:11]^2/sum(grwssn_avgs.pc$sdev^2)*100) %>%
  ggplot(aes(x=PC, y=percent_var)) +
  geom_col() +
  ggtitle("Percent Variance Explained")

#combine pcs with metadata
grwssn_avgs.pc.dat = data.frame(grwssn_avgs.pc$x)
grwssn_avgs_locs.pc = cbind(grwssn_avgs, grwssn_avgs.pc.dat)
grwssn_avgs_loadings = data.frame(varnames=rownames(grwssn_avgs.pc$rotation), grwssn_avgs.pc$rotation)
grwssn_avgs_loadings
```

#### Permanova 
```{r}
grwssn_avgs_locs.pc_dist <- grwssn_avgs_locs.pc %>% ungroup() %>% select(PC1:PC11)
dist_matrix_grwssn <- dist(grwssn_avgs_locs.pc_dist, method = "euclidian") #use a distance function to calculate euclidian distance in PCA space
permanova_results_grwssn <- adonis2(dist_matrix_grwssn ~ timeframe*elev_m*Lat, data = grwssn_avgs_locs.pc) #use adonis2 to run the permanova
permanova_results_grwssn #look at output 
#get stats per term in the model:
permanova_results_grwssn_terms <- adonis2(dist_matrix_grwssn ~ timeframe*elev_m*Lat, data = grwssn_avgs_locs.pc, by = "terms")
permanova_results_grwssn_terms

#LM on PCs (follow up on permanova)
lmer_results_grwssn <- grwssn_avgs_locs.pc %>%
  ungroup() %>% 
  select(timeframe, parent.pop, elev_m, Lat, Long, PC1:PC11) %>% 
  pivot_longer(starts_with("PC", ignore.case = FALSE), 
               names_to = "PC", values_to = "value") %>% 
  group_by(PC) %>% 
  nest(data=c(timeframe, parent.pop, elev_m, Lat, Long, value)) %>% 
  mutate(glm=map(data, ~ glm(value ~ timeframe*elev_m*Lat,
                               data=.x)),
         anova = map(glm, ~ broom.mixed::tidy(anova(.x))))

PC_anova_grwssn <- lmer_results_grwssn %>% select(-data, -glm) %>% unnest(anova) %>%
  select(PC, term, p.value) %>%
  filter(p.value < 0.05) %>%
  arrange(term, p.value)
PC_anova_grwssn #PC2 and PC4 most sig, PC7 also has sig timeframe effects 
#PC2 also has a significant timeframe*elev_m effect
#PC9 has sig timeframe*lat and timeframe*elev*lat effects 

lmer_results_grwssn %>% select(-data, -glm) %>% unnest(anova) %>% filter(PC=="PC1" | PC=="PC2" |PC=="PC3" | PC=="PC4") 
```

#### Plot climate change PCA
```{r}
#prep the pc for making a plot with arrows distinguishing recent vs. historical time 
grwssn_avgs_locs.pc_avg <- grwssn_avgs_locs.pc %>%
  group_by(parent.pop, elev_m, timeframe) %>%
  summarise(across(.cols=starts_with("PC", ignore.case = FALSE), .fns = mean)) %>%
  ungroup()

autoplot(grwssn_avgs.pc, data = grwssn_avgs,
         x=1, y=2,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

autoplot(grwssn_avgs.pc, data = grwssn_avgs,
         x=2, y=4,
         colour='elev_m', alpha=0.5,
         label=TRUE, label.label="parent.pop",
         loadings=TRUE, loadings.colour='black', loadings.linewidth = 0.7,
         loadings.label = TRUE, loadings.label.size=6, loadings.label.colour="black", 
         loadings.label.vjust = -0.2, loadings.label.repel=TRUE) +
   scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  theme_classic()

grwssn_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=timeframe, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

grwssn_avgs_locs.pc_avg %>% 
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC2, y=PC4, shape=timeframe, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=2, alpha=0.7) +
  geom_text_repel(aes(label = parent.pop)) +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed")  +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8)

#grwssn_avgs_locs.pc_avg %>% filter(parent.pop=="WL2")
#grwssn_avgs_loadings
```

```{r}
#PC1 = neg ann_tmean, tmean_wettest_month, ann_ppt
#PC2 = positive temp_seasonality and temp_ann_range, negative ppt
#PC4 = negative mean_diurnal_range
home_sites_pca_grwssn <- grwssn_avgs_locs.pc_avg %>%  
  mutate(group=str_c(parent.pop,elev_m))  %>%
  ggplot(aes(x=PC1, y=PC2, shape=timeframe, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  labs(x="PC1 (38.76%)", y="PC2 (26.73%)", color="Elevation (m)") +
  geom_vline(xintercept = 0, linetype="dashed") + geom_hline(yintercept = 0, linetype="dashed") +
  annotate("text", x = 0.73, y= -0.15, label = "WL2", colour = "purple", size=7) +
  geom_path(aes(group=group),arrow = arrow(length=unit(5, "points")), linewidth = .8) +
  annotate("text", x = -4, y = -5.2, label = "Warm \n Wet", size=6) +
  annotate("text", x = 3.5, y = -5.2, label = "Cold \n Dry", size=6) +
  annotate("text", x = -5.4, y = -3, label = "Low Temp \n Seasonality", size=6) +
  annotate("text", x = -5.4, y = 2.7, label = "High Temp \n Seasonality", size=6) +
  coord_cartesian(ylim = c(-4, 3), xlim = c(-4.5,4), clip = "off") +
  theme_classic() +
  theme(text=element_text(size=28))

## add WL2 garden 2023 and 2024
WL2GRDN_grwssn_pc_prep_2023 <- all_clim %>% 
  ungroup() %>% 
  filter(parent.pop=="WL2_Garden", year==2023, Season=="Growth Season") %>% 
  select(cwd:ppt, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_month, ann_ppt, ppt_seasonality, ppt_coldest_month) 
WL2GRDN_grwssn_predicted_2023 <- predict(grwssn_avgs.pc, newdata = WL2GRDN_grwssn_pc_prep_2023)

WL2GRDN_grwssn_pc_prep_2024 <- all_clim %>% 
  ungroup() %>% 
  filter(parent.pop=="WL2_Garden", year==2024, Season=="Growth Season") %>% #need 2024 Growth Season data for WL2 garden 
  select(cwd:ppt, ann_tmean, mean_diurnal_range, temp_seasonality, temp_ann_range, tmean_wettest_month, ann_ppt, ppt_seasonality, ppt_coldest_month) 
WL2GRDN_grwssn_2024_predicted <- predict(grwssn_avgs.pc, newdata = WL2GRDN_grwssn_pc_prep_2024)

home_sites_pca_grwssn$data <- rbind(home_sites_pca_grwssn$data, 
  data.frame(
    parent.pop = "WL2_Garden",
    elev_m = 2020,
    timeframe = c("2023", "2024"),
    PC1 = c(WL2GRDN_grwssn_predicted_2023[, "PC1"], WL2GRDN_grwssn_2024_predicted[, "PC1"]),
    PC2 = c(WL2GRDN_grwssn_predicted_2023[, "PC2"], WL2GRDN_grwssn_2024_predicted[, "PC2"]),
    PC3 = c(WL2GRDN_grwssn_predicted_2023[, "PC3"], WL2GRDN_grwssn_2024_predicted[, "PC3"]),
    PC4 = c(WL2GRDN_grwssn_predicted_2023[, "PC4"], WL2GRDN_grwssn_2024_predicted[, "PC4"]),
    PC5 = c(WL2GRDN_grwssn_predicted_2023[, "PC5"], WL2GRDN_grwssn_2024_predicted[, "PC5"]),
    PC6 = c(WL2GRDN_grwssn_predicted_2023[, "PC6"], WL2GRDN_grwssn_2024_predicted[, "PC6"]),
    PC7 = c(WL2GRDN_grwssn_predicted_2023[, "PC7"], WL2GRDN_grwssn_2024_predicted[, "PC7"]),
    PC8 = c(WL2GRDN_grwssn_predicted_2023[, "PC8"], WL2GRDN_grwssn_2024_predicted[, "PC8"]),
    PC9 = c(WL2GRDN_grwssn_predicted_2023[, "PC9"], WL2GRDN_grwssn_2024_predicted[, "PC9"]),
    PC10 = c(WL2GRDN_grwssn_predicted_2023[, "PC10"], WL2GRDN_grwssn_2024_predicted[, "PC10"]),
    PC11 = c(WL2GRDN_grwssn_predicted_2023[, "PC11"], WL2GRDN_grwssn_2024_predicted[, "PC11"]),
    group = c("new", "new2")
  )
)

home_sites_pca_grwssn + 
  geom_point(data=filter(home_sites_pca_grwssn$data, parent.pop == "WL2_Garden"), size=6, shape = 18, show.legend = FALSE) +
  annotate("text", x = -0.07, y= -2.6, label = "WL2 Garden \n 2023", colour = "purple", size=6) +
  annotate("text", x = 1.5, y= 3, label = "WL2 Garden 2024", colour = "purple", size=6) 
ggsave("../Figures/Growth_Season_PC1-PC2_PlusGarden.png", width = 7.4, height = 6, units = "in")
```


## WL2 Yearly Variation PCAs (Figure S2)
```{r}

```
