---
title: "Clim_dist_calcs"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code for calculating climate distance (Figure S3; Table S4)

## Load Libraries
```{r}
library(tidyverse)
library(boot) #for bootstrapping 
library(broom) #for tidy function
```

## Load Climate Data
```{r}
all_clim <- read_csv("../Processed.Data/Climate/All_Clim.csv")
```

## WL2 Garden Climate
```{r}
#Nov 2022-October 2023
wl2_garden_2023_wtryr <- all_clim %>% 
  filter(parent.pop=="WL2_Garden", Season=="Water Year", year==2023) %>%
  rename_with(.cols = c(8:22), ~paste0(.x, '_WL2')) %>% 
  select(cwd_WL2:ppt_coldest_month_WL2)
#Nov 2023-October 2024
wl2_garden_2024_wtryr <- all_clim %>% 
  filter(parent.pop=="WL2_Garden", Season=="Water Year", year==2024) %>%
  rename_with(.cols = c(8:22), ~paste0(.x, '_WL2')) %>% 
  select(cwd_WL2:ppt_coldest_month_WL2)
#July-December 2023 
wl2_garden_2023_grwssn <- all_clim %>% 
  filter(parent.pop=="WL2_Garden", Season=="Growth Season", year==2023) %>%
  rename_with(.cols = c(8:22), ~paste0(.x, '_WL2')) %>% 
  select(cwd_WL2:ppt_coldest_month_WL2)
#June-November 2024
wl2_garden_2024_grwssn <- all_clim %>% 
  filter(parent.pop=="WL2_Garden", Season=="Growth Season", year==2024) %>%
  rename_with(.cols = c(8:22), ~paste0(.x, '_WL2')) %>% 
  select(cwd_WL2:ppt_coldest_month_WL2)
```

## Gower's Climate Distance

(1/P) \* SUM ((absolute value(Ai - Bi)) / range(i)) for each variable

-   P = number of environmental variables = 13 (without CWD)

-   Ai = 30 year avg of that variable for the home site

-   Bi = July 2023-Dec 2023 avg of that variable for the WL2 garden

-   Range(i) = maximum - minimum of that variable in the whole data set
    (across sites)
    
### Prep the dataframes
```{r}
#pop info 
pops <- all_clim %>% 
  filter(!is.na(Lat), parent.pop!="WL2_Garden") %>% 
  select(parent.pop, elevation.group, elev_m, Lat, Long) %>% 
  unique()

#home climates
recent_wtryr_clim <- all_clim %>% 
  filter(timeframe=="recent", Season=="Water Year")
recent_wtryr_clim_nest <- recent_wtryr_clim %>% nest(.by=year) #nest to prepare for bootstrapping 
recent_wtryr_clim_nest

historic_wtryr_clim <- all_clim %>% 
  filter(timeframe=="historic", Season=="Water Year")
historic_wtryr_clim_nest <- historic_wtryr_clim %>% nest(.by=year) #nest to prepare for bootstrapping 
historic_wtryr_clim_nest

recent_grwssn_clim <- all_clim %>% 
  filter(timeframe=="recent", Season=="Growth Season", parent.pop!="WL2_Garden") 
recent_grwssn_clim_nest <- recent_grwssn_clim %>% nest(.by=year) #nest to prepare for bootstrapping 
recent_grwssn_clim_nest

historic_grwssn_clim <- all_clim %>% 
  filter(timeframe=="historic", Season=="Growth Season", parent.pop!="WL2_Garden")
historic_grwssn_clim_nest <- historic_grwssn_clim %>% nest(.by=year) #nest to prepare for bootstrapping 
historic_grwssn_clim_nest

#WL2 range prep
wl2_garden_2023_wtryr_range_prep <- wl2_garden_2023_wtryr %>% 
  rename_with(~str_remove(., "_WL2"), everything())
wl2_garden_2024_wtryr_range_prep <- wl2_garden_2024_wtryr %>% 
  rename_with(~str_remove(., "_WL2"), everything())
wl2_garden_2023_grwssn_range_prep <- wl2_garden_2023_grwssn %>% 
  rename_with(~str_remove(., "_WL2"), everything())
wl2_garden_2024_grwssn_range_prep <- wl2_garden_2024_grwssn %>% 
  rename_with(~str_remove(., "_WL2"), everything())
```

### Create the gower_calc function 

```{r}
#gower's calc function 
gowers_calc <- function(data, indices, P, garden_clim, garden_range) { #function with all of the code necessary for calculating gowers distance 
  #data = _clim_nest (recent or historical) - needs to be nested by year; P = # climate variables;
  #garden_clim = wl2 garden clim with _Wl2 at end of column names
  #garden_range = wl2 garden data with WL2 removed from the column names 
  
  data <-data[indices,] # subset per bootstrap indices
  
  data <- data %>% unnest(data) #unnest so the function can access the climate data
  
  data_means <- data %>% 
    group_by(parent.pop, elevation.group, elev_m, Lat, Long, timeframe) %>% 
    summarise_at(c("cwd",  "pck", "ppt", "tmn", "tmx", "ann_tmean", "mean_diurnal_range", 
                   "temp_seasonality", "temp_ann_range",
                 "tmean_wettest_month", "tmean_driest_month", "ann_ppt",
                 "ppt_seasonality","ppt_warmest_month", "ppt_coldest_month"),
               c(mean), na.rm = TRUE) #get 30 year averages for each climate variable 
  
  range_merge <- bind_rows(data_means, garden_range)
  
  WL2_home_climate_ranges <- range_merge %>% #calculate ranges
    ungroup() %>% 
  summarise(cwd_range=max(cwd)-min(cwd),
            pck_range=max(pck)-min(pck),
            ppt_range=max(ppt)-min(ppt), 
            tmn_range=max(tmn)-min(tmn), 
            tmx_range=max(tmx)-min(tmx), 
            ann_tmean_range=max(ann_tmean)-min(ann_tmean),
            mean_diurnal_range_range=max(mean_diurnal_range)-min(mean_diurnal_range),
            temp_seasonality_range=max(temp_seasonality)-min(temp_seasonality),
            temp_ann_range_range=max(temp_ann_range)-min(temp_ann_range),
            tmean_wettest_month_range=max(tmean_wettest_month)-min(tmean_wettest_month),
            tmean_driest_month_range=max(tmean_driest_month)-min(tmean_driest_month),
            ann_ppt_range=max(ann_ppt)-min(ann_ppt), 
            ppt_seasonality_range=max(ppt_seasonality)-min(ppt_seasonality),
            ppt_warmest_month_range=max(ppt_warmest_month)-min(ppt_warmest_month), 
            ppt_coldest_month_range=max(ppt_coldest_month)-min(ppt_coldest_month))
  
  WL2_home_climate <- bind_cols(garden_clim, data_means) #add WL2 climate data to home climate data 
  
  WL2_home_climate_with_ranges <- bind_cols(WL2_home_climate, WL2_home_climate_ranges) #add in ranges 
  
  gowers_calc_each_var <- WL2_home_climate_with_ranges %>% #variable by variable calc
  mutate(cwd_gowers=abs(cwd_WL2-cwd) / cwd_range,
         pck_gowers=abs(pck_WL2-pck) / pck_range,
         ppt_gowers=abs(ppt_WL2 - ppt) / ppt_range,
         tmn_gowers=abs(tmn_WL2 - tmn) / tmn_range,
         tmx_gowers=abs(tmx_WL2 - tmx) / tmx_range,
         ann_tmean_gowers=abs(ann_tmean_WL2 - ann_tmean) / ann_tmean_range,
         mean_diurnal_range_gowers=abs(mean_diurnal_range_WL2 - mean_diurnal_range) / mean_diurnal_range_range,
         temp_seasonality_gowers=abs(temp_seasonality_WL2 - temp_seasonality) / temp_seasonality_range,
         temp_ann_range_gowers=abs(temp_ann_range_WL2 - temp_ann_range) / temp_ann_range_range,
         tmean_wettest_month_gowers=abs(tmean_wettest_month_WL2 - tmean_wettest_month) / tmean_wettest_month_range,
         tmean_driest_month_gowers=abs(tmean_driest_month_WL2 - tmean_driest_month) / tmean_driest_month_range,
         ann_ppt_gowers=abs(ann_ppt_WL2 - ann_ppt) / ann_ppt_range,
         ppt_seasonality_gowers=abs(ppt_seasonality_WL2 - ppt_seasonality) / ppt_seasonality_range,
         ppt_warmest_month_gowers=abs(ppt_warmest_month_WL2 - ppt_warmest_month) / ppt_warmest_month_range,
         ppt_coldest_month_gowers=abs(ppt_coldest_month_WL2 - ppt_coldest_month) / ppt_coldest_month_range) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, ends_with("_gowers"))

 gowers_calc_per_pop <- gowers_calc_each_var %>% #final gowers calc 
  mutate(Gowers_Dist=(1/P)*(cwd_gowers + pck_gowers + ppt_gowers + tmn_gowers + tmx_gowers +
                                ann_tmean_gowers + mean_diurnal_range_gowers +
                                temp_seasonality_gowers +temp_ann_range_gowers +
                                tmean_wettest_month_gowers +
                                tmean_driest_month_gowers +ann_ppt_gowers +
                                ppt_seasonality_gowers + ppt_warmest_month_gowers +
                                ppt_coldest_month_gowers)) %>% 
  dplyr::select(parent.pop, elevation.group, elev_m, Gowers_Dist)
  
 gowers_calc_per_pop %>% pull(Gowers_Dist) #make the result a vector 
}
```

### Bootstrapping 

#### 2023 Garden
```{r}
# water year - recent
gowers.boot_2023_recent_wtryr <- boot(data=recent_wtryr_clim_nest, statistic=gowers_calc, R=1000, P=15, garden_clim=wl2_garden_2023_wtryr, garden_range=wl2_garden_2023_wtryr_range_prep) #will sample each row (year) with replacement 

for(i in 1:23) {
  plot(gowers.boot_2023_recent_wtryr, index=i) #distributions look normal for the most part 
}

boot_recent_wtryr_results_2023 <- tidy(gowers.boot_2023_recent_wtryr,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(timeframe="Recent", Season="Water Year", Year=2023)
boot_gowers_recent_wtryr_pops_2023 <- bind_cols(pops, boot_recent_wtryr_results_2023) %>% 
  select(parent.pop:Long, timeframe, Season, Gowers_Dist, conf.low, conf.high) %>% 
  arrange(Gowers_Dist)

# water year - historic
gowers.boot_2023_historic_wtryr <- boot(data=historic_wtryr_clim_nest, statistic=gowers_calc, R=1000, P=15, garden_clim=wl2_garden_2023_wtryr, garden_range=wl2_garden_2023_wtryr_range_prep) #will sample each row (year) with replacement 

for(i in 1:23) {
  plot(gowers.boot_2023_historic_wtryr, index=i) #distributions look normal for the most part 
}

boot_historic_wtryr_results_2023 <- tidy(gowers.boot_2023_historic_wtryr,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(timeframe="Historic", Season="Water Year", Year=2023)
boot_gowers_historic_wtryr_pops_2023 <- bind_cols(pops, boot_historic_wtryr_results_2023) %>% 
  select(parent.pop:Long, timeframe, Season, Gowers_Dist, conf.low, conf.high) %>% 
  arrange(Gowers_Dist)

# growth season - recent
gowers.boot_2023_recent_grwssn <- boot(data=recent_grwssn_clim_nest, statistic=gowers_calc, R=1000, P=15, garden_clim=wl2_garden_2023_grwssn, garden_range=wl2_garden_2023_grwssn_range_prep) #will sample each row (year) with replacement 

for(i in 1:23) {
  plot(gowers.boot_2023_recent_grwssn, index=i) #distributions look normal for the most part 
}

boot_recent_grwssn_results_2023 <- tidy(gowers.boot_2023_recent_grwssn,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(timeframe="Recent", Season="Growth Season", Year=2023)
boot_gowers_recent_grwssn_pops_2023 <- bind_cols(pops, boot_recent_grwssn_results_2023) %>% 
  select(parent.pop:Long, timeframe, Season, Gowers_Dist, conf.low, conf.high) %>% 
  arrange(Gowers_Dist)

# growth season - historic
gowers.boot_2023_historic_grwssn <- boot(data=historic_grwssn_clim_nest, statistic=gowers_calc, R=1000, P=15, garden_clim=wl2_garden_2023_grwssn, garden_range=wl2_garden_2023_grwssn_range_prep) #will sample each row (year) with replacement 

for(i in 1:23) {
  plot(gowers.boot_2023_historic_grwssn, index=i) #distributions look normal for the most part 
}

boot_historic_grwssn_results_2023 <- tidy(gowers.boot_2023_historic_grwssn,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(timeframe="Historic", Season="Growth Season", Year=2023)
boot_gowers_historic_grwssn_pops_2023 <- bind_cols(pops, boot_historic_grwssn_results_2023) %>% 
  select(parent.pop:Long, timeframe, Season, Gowers_Dist, conf.low, conf.high) %>% 
  arrange(Gowers_Dist)
```

#### 2024 Garden
```{r}
# water year - recent
gowers.boot_2024_recent_wtryr <- boot(data=recent_wtryr_clim_nest, statistic=gowers_calc, R=1000, P=15, garden_clim=wl2_garden_2024_wtryr, garden_range=wl2_garden_2024_wtryr_range_prep) #will sample each row (year) with replacement 

for(i in 1:23) {
  plot(gowers.boot_2024_recent_wtryr, index=i) #distributions look normal for the most part 
}

boot_recent_wtryr_results_2024 <- tidy(gowers.boot_2024_recent_wtryr,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(timeframe="Recent", Season="Water Year", Year=2024)
boot_gowers_recent_wtryr_pops_2024 <- bind_cols(pops, boot_recent_wtryr_results_2024) %>% 
  select(parent.pop:Long, timeframe, Season, Gowers_Dist, conf.low, conf.high) %>% 
  arrange(Gowers_Dist)

# water year - historic
gowers.boot_2024_historic_wtryr <- boot(data=historic_wtryr_clim_nest, statistic=gowers_calc, R=1000, P=15, garden_clim=wl2_garden_2024_wtryr, garden_range=wl2_garden_2024_wtryr_range_prep) #will sample each row (year) with replacement 

for(i in 1:23) {
  plot(gowers.boot_2024_historic_wtryr, index=i) #distributions look normal for the most part 
}

boot_historic_wtryr_results_2024 <- tidy(gowers.boot_2024_historic_wtryr,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(timeframe="Historic", Season="Water Year", Year=2024)
boot_gowers_historic_wtryr_pops_2024 <- bind_cols(pops, boot_historic_wtryr_results_2024) %>% 
  select(parent.pop:Long, timeframe, Season, Gowers_Dist, conf.low, conf.high) %>% 
  arrange(Gowers_Dist)

# growth season - recent
gowers.boot_2024_recent_grwssn <- boot(data=recent_grwssn_clim_nest, statistic=gowers_calc, R=1000, P=15, garden_clim=wl2_garden_2024_grwssn, garden_range=wl2_garden_2024_grwssn_range_prep) #will sample each row (year) with replacement 

for(i in 1:23) {
  plot(gowers.boot_2024_recent_grwssn, index=i) #distributions look normal for the most part 
}

boot_recent_grwssn_results_2024 <- tidy(gowers.boot_2024_recent_grwssn,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(timeframe="Recent", Season="Growth Season", Year=2024)
boot_gowers_recent_grwssn_pops_2024 <- bind_cols(pops, boot_recent_grwssn_results_2024) %>% 
  select(parent.pop:Long, timeframe, Season, Gowers_Dist, conf.low, conf.high) %>% 
  arrange(Gowers_Dist)

# growth season - historic
gowers.boot_2024_historic_grwssn <- boot(data=historic_grwssn_clim_nest, statistic=gowers_calc, R=1000, P=15, garden_clim=wl2_garden_2024_grwssn, garden_range=wl2_garden_2024_grwssn_range_prep) #will sample each row (year) with replacement 

for(i in 1:23) {
  plot(gowers.boot_2024_historic_grwssn, index=i) #distributions look normal for the most part 
}

boot_historic_grwssn_results_2024 <- tidy(gowers.boot_2024_historic_grwssn,conf.int=TRUE,conf.method="norm") %>%  #all pops
  rename(Gowers_Dist = statistic) %>% 
  mutate(timeframe="Historic", Season="Growth Season", Year=2024)
boot_gowers_historic_grwssn_pops_2024 <- bind_cols(pops, boot_historic_grwssn_results_2024) %>% 
  select(parent.pop:Long, timeframe, Season, Gowers_Dist, conf.low, conf.high) %>% 
  arrange(Gowers_Dist)
```

### Average gower's distance 
```{r}
wl2_gowers_avg <- bind_rows(boot_gowers_recent_wtryr_pops_2023,
                            boot_gowers_historic_wtryr_pops_2023,
                            boot_gowers_recent_grwssn_pops_2023,
                            boot_gowers_historic_grwssn_pops_2023,
                            boot_gowers_recent_wtryr_pops_2024,
                            boot_gowers_historic_wtryr_pops_2024,
                            boot_gowers_recent_grwssn_pops_2024,
                            boot_gowers_historic_grwssn_pops_2024) %>% 
  group_by(parent.pop, elevation.group, elev_m, Lat, Long, timeframe, Season) %>% 
  summarise(Avg_Gowers=mean(Gowers_Dist, na.rm=TRUE))
```

### Plot the results 
```{r}
# 2023 water year
boot_gowers_recent_wtryr_pops_2023 %>% 
  bind_rows(boot_gowers_historic_wtryr_pops_2023) %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_wrap(~timeframe)
ggsave("../Figures/wtryear_2023Gowers_fromWL2.png", width = 24, height = 8, units = "in")

# 2023 growth season
boot_gowers_recent_grwssn_pops_2023 %>% 
  bind_rows(boot_gowers_historic_grwssn_pops_2023) %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_wrap(~timeframe)
ggsave("../Figures/growthseason_2023Gowers_fromWL2.png", width = 24, height = 8, units = "in")

# 2024 water year 
boot_gowers_recent_wtryr_pops_2024 %>% 
  bind_rows(boot_gowers_historic_wtryr_pops_2024) %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_wrap(~timeframe)
ggsave("../Figures/wtryear_2024Gowers_fromWL2.png", width = 24, height = 8, units = "in")

# 2024 growth season
boot_gowers_recent_grwssn_pops_2024 %>% 
  bind_rows(boot_gowers_historic_grwssn_pops_2024) %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Gowers_Dist), y=Gowers_Dist, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high),
                width=.1, position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_wrap(~timeframe)
ggsave("../Figures/growthseason_2024Gowers_fromWL2.png", width = 24, height = 8, units = "in")

# average water year 
wl2_gowers_avg %>% 
  filter(Season=="Water Year") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Avg_Gowers), y=Avg_Gowers, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_wrap(~timeframe)
ggsave("../Figures/wtryear_AvgGowers_fromWL2.png", width = 24, height = 8, units = "in")

# average growth season 
wl2_gowers_avg %>% 
  filter(Season=="Growth Season") %>% 
  ggplot(aes(x=fct_reorder(parent.pop, Avg_Gowers), y=Avg_Gowers, group=parent.pop, fill=elev_m)) +
  geom_col(width = 0.7,position = position_dodge(0.75)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_gradient(low = "#F5A540", high = "#0043F0") +
  labs(y="Gowers Envtal Distance \n from WL2", fill="Elevation (m)", x="Population") +
  theme_classic() +
  theme(text=element_text(size=25), axis.text.x = element_text(angle = 45,  hjust = 1)) +
  facet_wrap(~timeframe)
ggsave("../Figures/growthseason_AvgGowers_fromWL2.png", width = 24, height = 8, units = "in")
```

## Subtraction Distances
```{r}

```

### Plot the results 
```{r}

```

## Merge all distances
```{r}
#add Elevation Distance and Geographic Distance 

```
