---
title: "Total_Fecundity"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code for Analyzing Total Fecundity 

Equation: p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) + p(Winter Surv)*p(Surv to Rep - y2)*Fruits(y2)

Calculated at the individual level 

## Libraries
```{r}
library(tidyverse)
library(lmerTest) #mixed models
```

## Read in Data
```{r}
## year 1 components 
wl2_establishment <- read_csv("../Processed.Data/WL2_Establishment.csv")
wl2_surv_to_rep_y1 <- read_csv("../Processed.Data/WL2_SurvtoRep_y1.csv")
wl2_fruits_y1 <- read_csv("../Processed.Data/WL2_Fruits_Y1.csv")

## year 2 components 
winter_surv <- read_csv("../Processed.Data/WL2_WinterSurv.csv")
wl2_surv_to_rep_y2 <- read_csv("../Processed.Data/WL2_Surv_to_Rep_Y2.csv")
wl2_fruits_y2 <- read_csv("../Processed.Data/WL2_Fruits_Y2.csv")
```

## Calculate Total Fecundity 
```{r}
wl2_total_fecundity <- left_join(wl2_establishment, wl2_surv_to_rep_y1) %>% 
  left_join(wl2_fruits_y1) %>% 
  select(-death.date, -bud.date) %>% #remove y1 death and bud dates 
  mutate(mf=as.numeric(mf), rep=as.numeric(rep)) %>% #convert these cols to numeric for merging 
  left_join(winter_surv) %>% 
  select(-death.date) %>% 
  left_join(wl2_surv_to_rep_y2) %>% 
  left_join(wl2_fruits_y2) %>% 
  select(-death.date) %>% 
  mutate(SurvtoRep_Y1=if_else(is.na(SurvtoRep_Y1), 0, SurvtoRep_Y1),
         y1_fruits=if_else(is.na(y1_fruits), 0, y1_fruits),
         y2_fruits=if_else(is.na(y2_fruits), 0, y2_fruits),
         WinterSurv=if_else(is.na(WinterSurv), 0, WinterSurv),
         SurvtoRep_y2=if_else(is.na(SurvtoRep_y2), 0, SurvtoRep_y2)) %>% 
  mutate(Total_Fecundity=(Establishment*SurvtoRep_Y1*y1_fruits) + (WinterSurv*SurvtoRep_y2*y2_fruits))

wl2_total_fecundity %>% 
  filter(Total_Fecundity>0) %>% 
  group_by(pop) %>% 
  summarise(n=n()) %>% 
  arrange(n) #SQ1 and WR only have 1 indiv 
```

## Check Distribution
```{r}
wl2_total_fecundity %>% 
  ggplot(aes(x=Total_Fecundity)) +
  geom_histogram()
```

## Try transformations
```{r}
wl2_total_fecundity_transformations <-
  wl2_total_fecundity %>% 
  mutate(logTotalFecundity=log(Total_Fecundity+1),
         log10TotalFecundity=log10(Total_Fecundity+1))

wl2_total_fecundity_transformations %>% 
  ggplot(aes(x=logTotalFecundity)) +
  geom_histogram()

wl2_total_fecundity_transformations %>% 
  ggplot(aes(x=log10TotalFecundity)) +
  geom_histogram()

## didn't help 
```

## Scale predictors
```{r}
wl2_total_fecundity_scaled <- wl2_total_fecundity %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  rename(AvgGD_Recent_Wtr_Year=`AvgGD_Recent_Water Year`,
         AvgGD_Recent_GrwSsn=`AvgGD_Recent_Growth Season`,
         AvgGD_Historic_Wtr_Year=`AvgGD_Historic_Water Year`,
         AvgGD_Historic_GrwSsn=`AvgGD_Historic_Growth Season`) %>% 
  mutate_at(c("AvgGD_Historic_GrwSsn", "AvgGD_Historic_Wtr_Year",
              "AvgGD_Recent_GrwSsn", "AvgGD_Recent_Wtr_Year",
              "AvgTempDist_Historic_GrwSsn", "AvgTempDist_Historic_Wtr_Year",
              "AvgTempDist_Recent_GrwSsn", "AvgTempDist_Recent_Wtr_Year",
              "AvgPPTDist_Historic_GrwSsn", "AvgPPTDist_Historic_Wtr_Year",
              "AvgPPTDist_Recent_GrwSsn", "AvgPPTDist_Recent_Wtr_Year",
              "Geographic_Dist"), scale)
```

## Plot group means and variances to facilitate choosing the right error distribution
```{r}
wl2_total_fecundity %>% 
  group_by(pop) %>% 
  summarise(GroupMeans=mean(Total_Fecundity), GroupVars=var(Total_Fecundity)) %>% 
  ggplot(aes(x=GroupMeans, y=GroupVars)) +
  geom_point() +
  geom_abline(intercept = 0, slope=1, colour="blue")
```
Based on this website, quasipoisson might work. https://r.qcbs.ca/workshop07/book-en/choose-an-error-distribution.html 

## Try models with alternative distributions 
Poisson, negative binomial, log-normal, beta 
```{r}
total_fecund_poisson <- glmer(Total_Fecundity ~ (1|pop/mf) + (1|block), 
                              family = poisson,
                              data=wl2_total_fecundity_scaled)
summary(total_fecund_poisson)

plot(total_fecund_poisson, which = 1) #residuals plot 
#doesn't look great

total_fecund_quasipoisson <- glmer(Total_Fecundity ~ (1|pop/mf) + (1|block), 
                              family = quasipoisson,
                              data=wl2_total_fecundity_scaled)
#Error: "quasi" families cannot be used in glmer
```

