---
title: "Total_Fecundity"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code for Analyzing Total Fecundity 

Equation: p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) + p(Winter Surv)*p(Surv to Rep - y2)*Fruits(y2)

Calculated at the individual level 

## Libraries
```{r}
library(tidyverse)
library(lmerTest) #mixed models
library(broom.mixed) #tidy model
library(emmeans) #for pairwise tests
library(performance) #for calculating rsq 

sem <- function(x, na.rm=FALSE) {           #for calculating standard error
  sd(x,na.rm=na.rm)/sqrt(length(na.omit(x)))
} 
```

## Read in Fitness Data
```{r}
## year 1 components 
wl2_establishment <- read_csv("../Processed.Data/WL2_Establishment.csv")
wl2_surv_to_rep_y1 <- read_csv("../Processed.Data/WL2_SurvtoRep_y1.csv")
wl2_fruits_y1 <- read_csv("../Processed.Data/WL2_Fruits_Y1.csv")

## year 2 components 
winter_surv <- read_csv("../Processed.Data/WL2_WinterSurv.csv")
wl2_surv_to_rep_y2 <- read_csv("../Processed.Data/WL2_Surv_to_Rep_Y2.csv")
wl2_fruits_y2 <- read_csv("../Processed.Data/WL2_Fruits_Y2.csv")
```

## Read in Size Pre-Transplant 
```{r}
wl2_pretrans_size1 <- read_csv("../Raw.Data/WL2_DNA_Collection_Size_survey_combined20230703_corrected.csv")

wl2_pretrans_size2 <- read_csv("../Raw.Data/WL2_Extras_DNA_collection_size_survey_combined20230706_corrected.csv") %>% 
  filter(!is.na(`height (cm)`)) #to get rid of genotypes that were measured on the other data sheet (NAs on this sheet)

wl2_pretrans_size_all <- bind_rows(wl2_pretrans_size1, wl2_pretrans_size2) %>%
  rename(pop=Pop, height.cm = `height (cm)`, long.leaf.cm=`longest leaf (cm)`) %>% 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  select(-DNA)
head(wl2_pretrans_size_all)
```

## Calculate Total Fecundity 
```{r}
wl2_total_fecundity <- left_join(wl2_establishment, wl2_surv_to_rep_y1) %>% 
  left_join(wl2_fruits_y1) %>% 
  select(-death.date, -bud.date) %>% #remove y1 death and bud dates 
  mutate(mf=as.numeric(mf), rep=as.numeric(rep)) %>% #convert these cols to numeric for merging 
  left_join(winter_surv) %>% 
  select(-death.date) %>% 
  left_join(wl2_surv_to_rep_y2) %>% 
  left_join(wl2_fruits_y2) %>% 
  select(-death.date) %>% 
  mutate(SurvtoRep_Y1=if_else(is.na(SurvtoRep_Y1), 0, SurvtoRep_Y1),
         y1_fruits=if_else(is.na(y1_fruits), 0, y1_fruits),
         y2_fruits=if_else(is.na(y2_fruits), 0, y2_fruits),
         WinterSurv=if_else(is.na(WinterSurv), 0, WinterSurv),
         SurvtoRep_y2=if_else(is.na(SurvtoRep_y2), 0, SurvtoRep_y2)) %>% 
  mutate(Total_Fecundity=(Establishment*SurvtoRep_Y1*y1_fruits) + (WinterSurv*SurvtoRep_y2*y2_fruits)) %>% 
  left_join(wl2_pretrans_size_all)

wl2_total_fecundity %>% 
  filter(Total_Fecundity>0) %>% 
  group_by(pop) %>% 
  summarise(n=n()) %>% 
  arrange(n) #SQ1 and WR only have 1 indiv 
```

## Check Distribution
```{r}
wl2_total_fecundity %>% 
  ggplot(aes(x=Total_Fecundity)) +
  geom_histogram()
```

## Try transformations
```{r}
wl2_total_fecundity_transformations <-
  wl2_total_fecundity %>% 
  mutate(logTotalFecundity=log(Total_Fecundity+1),
         log10TotalFecundity=log10(Total_Fecundity+1))

wl2_total_fecundity_transformations %>% 
  ggplot(aes(x=logTotalFecundity)) +
  geom_histogram()

wl2_total_fecundity_transformations %>% 
  ggplot(aes(x=log10TotalFecundity)) +
  geom_histogram()

## didn't help 
```

## Scale predictors
```{r}
wl2_total_fecundity_scaled <- wl2_total_fecundity %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  rename(AvgGD_Recent_Wtr_Year=`AvgGD_Recent_Water Year`,
         AvgGD_Recent_GrwSsn=`AvgGD_Recent_Growth Season`,
         AvgGD_Historic_Wtr_Year=`AvgGD_Historic_Water Year`,
         AvgGD_Historic_GrwSsn=`AvgGD_Historic_Growth Season`) %>% 
  mutate_at(c("AvgGD_Historic_GrwSsn", "AvgGD_Historic_Wtr_Year",
              "AvgGD_Recent_GrwSsn", "AvgGD_Recent_Wtr_Year",
              "AvgTempDist_Historic_GrwSsn", "AvgTempDist_Historic_Wtr_Year",
              "AvgTempDist_Recent_GrwSsn", "AvgTempDist_Recent_Wtr_Year",
              "AvgPPTDist_Historic_GrwSsn", "AvgPPTDist_Historic_Wtr_Year",
              "AvgPPTDist_Recent_GrwSsn", "AvgPPTDist_Recent_Wtr_Year",
              "Geographic_Dist"), scale)
```

## Plot group means and variances to facilitate choosing the right error distribution
```{r}
wl2_total_fecundity %>% 
  group_by(pop) %>% 
  summarise(GroupMeans=mean(Total_Fecundity), GroupVars=var(Total_Fecundity)) %>% 
  ggplot(aes(x=GroupMeans, y=GroupVars)) +
  geom_point() +
  geom_abline(intercept = 0, slope=1, colour="blue")
```
Based on this website, quasipoisson might work. https://r.qcbs.ca/workshop07/book-en/choose-an-error-distribution.html 

## Try models with poisson and quasi-poisson 

### Basic model - Poisson and Quasi-Poisson 
```{r}
total_fecund_poisson <- glmer(Total_Fecundity ~ (1|pop/mf) + (1|block), 
                              family = poisson,
                              data=wl2_total_fecundity_scaled)
summary(total_fecund_poisson)

plot(total_fecund_poisson, which = 1) #residuals plot 
#doesn't look great
deviance(total_fecund_poisson)/df.residual(total_fecund_poisson) #check for overdispersion (caused by excess of 0s)

##try quasi-poisson model to deal with the above (and based on the above plot)
##Can't directly use quasi families with glmer so have to use Ben Bolker's quasi-likelihood adjustment function:
#https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#fitting-models-with-overdispersion
#also got help from: https://stackoverflow.com/questions/68915173/how-do-i-fit-a-quasi-poisson-model-with-lme4-or-glmmtmb
quasi_table <- function(model,ctab=coef(summary(model))) {
    phi <- sum(residuals(model, type="pearson")^2)/df.residual(model) #overdispersion parameter
    qctab <- within(as.data.frame(ctab),
    {   `Std. Error` <- `Std. Error`*sqrt(phi)
        `z value` <- Estimate/`Std. Error`
        `Pr(>|z|)` <- 2*pnorm(abs(`z value`), lower.tail=FALSE)
    })
    return(qctab)
}
printCoefmat(quasi_table(total_fecund_poisson),digits=2)
#As specified, the estimates are identical; the standard errors and p-values have been appropriately inflated, the z-values have been appropriately deflated.
```

### Add in Gowers climate distance
```{r}
total_fecund_models_CD_GD <- tribble(
  ~name,          ~f,
  "1_pop.block",      "Total_Fecundity ~  height.cm + (1|pop/mf) + (1|block)", 
  "2_WY_Recent",      "Total_Fecundity ~  AvgGD_Recent_Wtr_Year + Geographic_Dist + height.cm + (1|pop/mf) + (1|block)",
  "3_WY_Historical",  "Total_Fecundity ~  AvgGD_Historic_Wtr_Year + Geographic_Dist + height.cm + (1|pop/mf) + (1|block)",
  "4_GS_Recent",      "Total_Fecundity ~  AvgGD_Recent_GrwSsn + Geographic_Dist + height.cm + (1|pop/mf) + (1|block)", 
  "5_GS_Historical",  "Total_Fecundity ~  AvgGD_Historic_GrwSsn + Geographic_Dist + height.cm + (1|pop/mf) + (1|block)"
)
#run the models 
total_fecund_models_CD_GD <- total_fecund_models_CD_GD %>%
  mutate(glmer = map(f, ~ glmer(as.formula(.), 
                            family = poisson,
                            data = wl2_total_fecundity_scaled)))

#show results:
total_fecund_models_CD_GD %>% mutate(tidy=map(glmer, tidy)) %>% unnest(tidy) %>%
  select(name, term:p.value) %>% 
  filter(str_detect(term, "GD") | term=="Geographic_Dist" | term=="height.cm") %>%
  drop_na(p.value)

#quasi-poisson correction:
phi_function <- function(model) {
    phi <- sum(residuals(model, type="pearson")^2)/df.residual(model) 
    return(phi)
} #function for calculating the overdispersion paramter

total_fecund_models_CD_GD %>% 
  mutate(phi=map(glmer, phi_function)) %>% 
  mutate(tidy=map(glmer, tidy)) %>% 
  unnest(tidy) %>% 
  unnest(phi) %>% 
  transmute(term = term, estimate = estimate, 
                std.error = std.error * sqrt(phi),
                statistic = estimate/std.error,
                p.value = 2*pnorm(abs(statistic), lower.tail=FALSE)) %>% 
  filter(str_detect(term, "GD") | term=="Geographic_Dist" | term=="height.cm") %>%
  drop_na(p.value) #after correction, no clim dist sig

total_fecund_models_CD_GD %>% mutate(rsq=map(glmer, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-f:-glmer)

#pairwise test between WL2 and BH (the pop with the highest total fecundity):
wtr_yr_recent_mod <- glmer(Total_Fecundity ~  AvgGD_Recent_Wtr_Year + Geographic_Dist + height.cm + (1|pop/mf) + (1|block), family = poisson, data = wl2_total_fecundity_scaled)
#get min and max GD
min(wl2_total_fecundity_scaled$AvgGD_Recent_Wtr_Year) 
max(wl2_total_fecundity_scaled$AvgGD_Recent_Wtr_Year)
minmaxGD = list(AvgGD_Recent_Wtr_Year = c(-1.358033, 2.153113))

# Calculate EMMs at the min and max GD
emm_results <- emmeans(wtr_yr_recent_mod, ~ AvgGD_Recent_Wtr_Year, at = minmaxGD)
# View the results
summary(emm_results)

# Perform pairwise comparison
pairs(emm_results) 
#P=0.0099
#but this is based off the poisson model, not the quasi-poisson...
```

### Temp/ppt dist
```{r}
total_fecund_SUB_models_CD_GD <- tribble(
  ~name,          ~f,
  "1_pop.block",      "Total_Fecundity ~  height.cm + (1|pop/mf) + (1|block)", 
  "2_WY_Recent",      "Total_Fecundity ~  AvgTempDist_Recent_Wtr_Year + AvgPPTDist_Recent_Wtr_Year + Geographic_Dist + height.cm + (1|pop/mf) + (1|block)",
  "3_WY_Historical",  "Total_Fecundity ~  AvgTempDist_Historic_Wtr_Year + AvgPPTDist_Historic_Wtr_Year + Geographic_Dist + height.cm + (1|pop/mf) + (1|block)",
  "4_GS_Recent",      "Total_Fecundity ~  AvgTempDist_Recent_GrwSsn + AvgPPTDist_Recent_GrwSsn + Geographic_Dist + height.cm + (1|pop/mf) + (1|block)", 
  "5_GS_Historical",  "Total_Fecundity ~  AvgTempDist_Historic_GrwSsn + AvgPPTDist_Historic_GrwSsn + Geographic_Dist + height.cm + (1|pop/mf) + (1|block)"
)

#run the models 
total_fecund_SUB_models_CD_GD <- total_fecund_SUB_models_CD_GD %>%
  mutate(glmer = map(f, ~ glmer(as.formula(.), 
                            family = poisson,
                            data = wl2_total_fecundity_scaled)))
#convergence warning for water year recent model 

#show results:
total_fecund_SUB_models_CD_GD %>% mutate(tidy=map(glmer, tidy)) %>% unnest(tidy) %>%
  select(name, term:p.value) %>% 
  filter(str_detect(term, "GD") | term=="Geographic_Dist" | term=="height.cm") %>%
  drop_na(p.value)

#quasi-poisson correction:
total_fecund_SUB_models_CD_GD %>% 
  mutate(phi=map(glmer, phi_function)) %>% 
  mutate(tidy=map(glmer, tidy)) %>% 
  unnest(tidy) %>% 
  unnest(phi) %>% 
  transmute(term = term, estimate = estimate,
                std.error = std.error * sqrt(phi),
                statistic = estimate/std.error,
                p.value = 2*pnorm(abs(statistic), lower.tail=FALSE)) %>% 
  filter(str_detect(term, "GD") | term=="Geographic_Dist" | term=="height.cm") %>%
  drop_na(p.value) 

total_fecund_SUB_models_CD_GD %>% mutate(rsq=map(glmer, r2)) %>% 
  unnest(rsq) %>% 
  unnest(rsq) %>% 
  select(-f:-glmer)

#investigate water year recent model:
mod_test <- glmer(Total_Fecundity ~ AvgTempDist_Recent_Wtr_Year + AvgPPTDist_Recent_Wtr_Year + Geographic_Dist + (1|pop/mf) + (1|block), 
                              family = poisson,
                              data=wl2_total_fecundity_scaled)
summary(mod_test)
#Warning: Model failed to converge with max|grad| = 0.121913 (tol = 0.002, component 1)Warning: Model is nearly unidentifiable: very large eigenvalue
 #- Rescale variables?
```

## Figures
```{r}
total_fecund_summary <- wl2_total_fecundity %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  rename(AvgGD_Recent_Water_Yr=`AvgGD_Recent_Water Year`) %>% 
  group_by(pop, elev_m, AvgGD_Recent_Water_Yr) %>% 
  summarise(meanTotalFecund=mean(Total_Fecundity, na.rm=TRUE), semEst=sem(Total_Fecundity, na.rm=TRUE))
total_fecund_summary %>% arrange(meanTotalFecund) #BH had the highest total fecundity 

total_fecund_summary %>% 
  ggplot(aes(x=AvgGD_Recent_Water_Yr, y=meanTotalFecund, group = pop, color=elev_m)) +
  scale_colour_gradient(low = "#F5A540", high = "#0043F0") +
  geom_point(size=6) + 
  geom_errorbar(aes(ymin=meanTotalFecund-semEst,ymax=meanTotalFecund+semEst),width=.02,linewidth = 2) +
  theme_classic() + 
  scale_y_continuous(expand = c(0.01, 0)) +
  labs(y="Total Fecundity", x="Recent Water Year Climate Dist", 
       color="Elevation (m)") +
  annotate("text", x = 0.2532683, y= 2.4, label = "WL2", 
           colour = "purple", fontface="bold", size = 22 / .pt) +
  theme(text=element_text(size=30))
ggsave("../Figures/Indiv_Fit_Comps/TotalFecund_GD_Recent_WtrYr.png",
       width = 12, height = 6, units = "in")
```
