---
title: "Total_Fitness"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code for Analyzing Total Fitness (Tables 2-3, S4, S5)

Equation: p(Establishment)*p(Surv to Rep - y1)*Fruits(y1) + p(Winter Surv)*p(Surv to Rep - y2)*Fruits(y2)

Calculated at the individual level 

## Libraries
```{r}
library(tidyverse)
library(broom.mixed) #tidy models
library(performance) #for calculating rsq 
library(glmmTMB) #for zero-inflated models 
library(bbmle) #for AICtab 
```

## Read in Fitness Data
```{r}
## year 1 components 
wl2_establishment <- read_csv("../Processed.Data/WL2_Establishment.csv")
wl2_surv_to_rep_y1 <- read_csv("../Processed.Data/WL2_SurvtoRep_y1.csv")
wl2_fruits_y1 <- read_csv("../Processed.Data/WL2_Fruits_Y1.csv")

## year 2 components 
winter_surv <- read_csv("../Processed.Data/WL2_WinterSurv.csv")
wl2_surv_to_rep_y2 <- read_csv("../Processed.Data/WL2_Surv_to_Rep_Y2.csv")
wl2_fruits_y2 <- read_csv("../Processed.Data/WL2_Fruits_Y2.csv")
```

## Read in Size Pre-Transplant 
```{r}
wl2_pretrans_size1 <- read_csv("../Raw.Data/WL2_DNA_Collection_Size_survey_combined20230703_corrected.csv")

wl2_pretrans_size2 <- read_csv("../Raw.Data/WL2_Extras_DNA_collection_size_survey_combined20230706_corrected.csv") %>% 
  filter(!is.na(`height (cm)`)) #to get rid of genotypes that were measured on the other data sheet (NAs on this sheet)

wl2_pretrans_size_all <- bind_rows(wl2_pretrans_size1, wl2_pretrans_size2) %>%
  rename(pop=Pop, height.cm = `height (cm)`, long.leaf.cm=`longest leaf (cm)`) %>% 
  unite(Genotype, pop:rep, sep="_", remove = FALSE) %>% 
  select(-DNA)
head(wl2_pretrans_size_all)
```

## Calculate Total Fitness 
```{r}
wl2_total_fitness <- left_join(wl2_establishment, wl2_surv_to_rep_y1) %>% 
  left_join(wl2_fruits_y1) %>% 
  select(-death.date, -bud.date) %>% #remove y1 death and bud dates 
  mutate(mf=as.numeric(mf), rep=as.numeric(rep)) %>% #convert these cols to numeric for merging 
  left_join(winter_surv) %>% 
  select(-death.date) %>% 
  left_join(wl2_surv_to_rep_y2) %>% 
  left_join(wl2_fruits_y2) %>% 
  select(-death.date) %>% 
  mutate(SurvtoRep_Y1=if_else(is.na(SurvtoRep_Y1), 0, SurvtoRep_Y1),
         y1_fruits=if_else(is.na(y1_fruits), 0, y1_fruits),
         y2_fruits=if_else(is.na(y2_fruits), 0, y2_fruits),
         WinterSurv=if_else(is.na(WinterSurv), 0, WinterSurv),
         SurvtoRep_y2=if_else(is.na(SurvtoRep_y2), 0, SurvtoRep_y2)) %>% 
  mutate(Total_Fitness=(Establishment*SurvtoRep_Y1*y1_fruits) + (WinterSurv*SurvtoRep_y2*y2_fruits)) %>% 
  left_join(wl2_pretrans_size_all)
write_csv(wl2_total_fitness, "../Processed.Data/WL2_Total_Fitness.csv")

wl2_total_fitness %>% 
  filter(Total_Fitness>0) %>% 
  group_by(pop) %>% 
  summarise(n=n()) %>% 
  arrange(n) #SQ1 and WR only have 1 indiv 
```

## Check Distribution
```{r}
wl2_total_fitness %>% 
  ggplot(aes(x=Total_Fitness)) +
  geom_histogram()
#Highly zero-inflated
```

## Try transformations
```{r}
wl2_total_fitness_transformations <-
  wl2_total_fitness %>% 
  mutate(logTotalFitness=log(Total_Fitness+1),
         log10TotalFitness=log10(Total_Fitness+1))

wl2_total_fitness_transformations %>% 
  ggplot(aes(x=logTotalFitness)) +
  geom_histogram()

wl2_total_fitness_transformations %>% 
  ggplot(aes(x=log10TotalFitness)) +
  geom_histogram()

## didn't help 
```

## Scale predictors
```{r}
wl2_total_fitness_scaled <- wl2_total_fitness %>% 
  filter(pop!="SQ1", pop!="WR") %>% 
  rename(AvgGD_Recent_Wtr_Year=`AvgGD_Recent_Water Year`,
         AvgGD_Recent_GrwSsn=`AvgGD_Recent_Growth Season`,
         AvgGD_Historic_Wtr_Year=`AvgGD_Historic_Water Year`,
         AvgGD_Historic_GrwSsn=`AvgGD_Historic_Growth Season`) %>% 
  mutate(across(c("AvgGD_Historic_GrwSsn", "AvgGD_Historic_Wtr_Year",
              "AvgGD_Recent_GrwSsn", "AvgGD_Recent_Wtr_Year",
              "AvgTempDist_Historic_GrwSsn", "AvgTempDist_Historic_Wtr_Year",
              "AvgTempDist_Recent_GrwSsn", "AvgTempDist_Recent_Wtr_Year",
              "AvgPPTDist_Historic_GrwSsn", "AvgPPTDist_Historic_Wtr_Year",
              "AvgPPTDist_Recent_GrwSsn", "AvgPPTDist_Recent_Wtr_Year",
              "Geographic_Dist", "height.cm"), scale, .names = "{.col}_scaled")) %>% 
  arrange(Total_Fitness)
```

## Geo fixed for Clim Dist Predictions 
```{r}
geo_fixed <- wl2_total_fitness_scaled %>% 
  summarise(Geographic_Dist=mean(Geographic_Dist),
            Geographic_Dist_scaled=mean(Geographic_Dist_scaled))

pred_df_gowers <- tibble(AvgGD_Recent_Wtr_Year = seq(from = 0.25, to=0.64, length.out=46)) %>% 
  mutate(AvgGD_Recent_Wtr_Year_scaled=scale(AvgGD_Recent_Wtr_Year)) %>% 
  bind_cols(geo_fixed)
```

## Basic model
Use glmmTMB to try different zero-inflated models 
```{r}
total_fit_zero_poisson <- glmmTMB(Total_Fitness ~ height.cm_scaled + (1|pop/mf) + (1|block),
                                     data=wl2_total_fitness_scaled,
                                     ziformula=~1, #zeroinflation parameter applying to all observations
                                     family=poisson)

total_fit_zero_negbinom <- update(total_fit_zero_poisson, family=nbinom2)

total_fit_zero_negbinom1 <- update(total_fit_zero_poisson, family=nbinom1)

AICctab(total_fit_zero_poisson, total_fit_zero_negbinom, total_fit_zero_negbinom1)
#best fit is negbinom1

#diagnostics 
deviance(total_fit_zero_negbinom1)/df.residual(total_fit_zero_negbinom1) #check for overdispersion - less than 1 which is good 
plot(predict(total_fit_zero_negbinom1, type="response"), residuals(total_fit_zero_negbinom1),
     main="glmmTMB-negbinom", xlab="Fitted", ylab="Residuals") #residuals plot
#not perfect, but probably fine?
plot(residuals(total_fit_zero_negbinom1)) #index plot 

plot(predict(total_fit_zero_poisson, type="response"), residuals(total_fit_zero_poisson),
     main="glmmTMB-poisson", xlab="Fitted", ylab="Residuals") 
deviance(total_fit_zero_poisson)/df.residual(total_fit_zero_poisson)

plot(predict(total_fit_zero_negbinom, type="response"), residuals(total_fit_zero_negbinom), xlab="Fitted", ylab="Residuals") 
deviance(total_fit_zero_negbinom)/df.residual(total_fit_zero_negbinom)

#results
summary(total_fit_zero_negbinom1) #mf explains very little variation
check_singularity(total_fit_zero_negbinom1)
r2(total_fit_zero_negbinom1, tolerance = 1e-09) #use marginal r2 because of singularity 
```

## Temp/ppt dist
```{r}
total_fit_SUB_CD_models <- tribble(
  ~name,          ~f,
  "1_pop.block",      "Total_Fitness ~  height.cm_scaled + (1|pop) + (1|block)", 
  "2_WY_Recent",      "Total_Fitness ~  AvgTempDist_Recent_Wtr_Year_scaled + AvgPPTDist_Recent_Wtr_Year_scaled + Geographic_Dist_scaled + (1|pop) + (1|block)",
  "3_WY_Historical",  "Total_Fitness ~  AvgTempDist_Historic_Wtr_Year_scaled + AvgPPTDist_Historic_Wtr_Year_scaled + Geographic_Dist_scaled + (1|pop) + (1|block)",
  "4_GS_Recent",      "Total_Fitness ~  AvgTempDist_Recent_GrwSsn_scaled + AvgPPTDist_Recent_GrwSsn_scaled + Geographic_Dist_scaled + (1|pop) + (1|block)", 
  "5_GS_Historical",  "Total_Fitness ~  AvgTempDist_Historic_GrwSsn_scaled + AvgPPTDist_Historic_GrwSsn_scaled + Geographic_Dist_scaled + (1|pop) + (1|block)",
  "6_WY_Recent_height",      "Total_Fitness ~  AvgTempDist_Recent_Wtr_Year_scaled + AvgPPTDist_Recent_Wtr_Year_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)",
  "7_WY_Historical_height",  "Total_Fitness ~  AvgTempDist_Historic_Wtr_Year_scaled + AvgPPTDist_Historic_Wtr_Year_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)" ,
  "8_GS_Recent_height",      "Total_Fitness ~  AvgTempDist_Recent_GrwSsn_scaled + AvgPPTDist_Recent_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)", 
  "9_GS_Historical_height",  "Total_Fitness ~  AvgTempDist_Historic_GrwSsn_scaled + AvgPPTDist_Historic_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)"
)

#run the models 
total_fit_SUB_CD_models <- total_fit_SUB_CD_models %>%
  mutate(glmmTMB = map(f, ~ glmmTMB(as.formula(.), 
                            family = nbinom1, ziformula=~1,
                            data = wl2_total_fitness_scaled)))

#show results:
total_fit_SUB_CD_models %>% mutate(tidy=map(glmmTMB, tidy)) %>% unnest(tidy) %>%
  select(name, term:p.value) %>% 
  filter(str_detect(term, "scaled")) %>%
  drop_na(p.value)

total_fit_SUB_CD_models %>% 
  mutate(rsq=map(glmmTMB, r2),
         singularity=map(glmmTMB, check_singularity)) %>% 
  unnest(c(rsq, singularity)) %>% 
  unnest(rsq) %>% 
  select(-f:-glmmTMB)
```

## Gowers dist
```{r}
total_fit_models_CD_GD <- tribble(
  ~name,          ~f,
  "1_pop.block",      "Total_Fitness ~  (1|pop) + (1|block)", 
  "2_WY_Recent",      "Total_Fitness ~  AvgGD_Recent_Wtr_Year_scaled + Geographic_Dist_scaled + (1|pop) + (1|block)",
  "3_WY_Historical",  "Total_Fitness ~  AvgGD_Historic_Wtr_Year_scaled + Geographic_Dist_scaled + (1|pop) + (1|block)",
  "4_GS_Recent",      "Total_Fitness ~  AvgGD_Recent_GrwSsn_scaled + Geographic_Dist_scaled + (1|pop) + (1|block)", 
  "5_GS_Historical",  "Total_Fitness ~  AvgGD_Historic_GrwSsn_scaled + Geographic_Dist_scaled + (1|pop) + (1|block)",
  "6_WY_Recent_height",      "Total_Fitness ~  AvgGD_Recent_Wtr_Year_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)",
  "7_WY_Historical_height",  "Total_Fitness ~  AvgGD_Historic_Wtr_Year_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)",
  "8_GS_Recent_height",      "Total_Fitness ~  AvgGD_Recent_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)", 
  "9_GS_Historical_height",  "Total_Fitness ~  AvgGD_Historic_GrwSsn_scaled + Geographic_Dist_scaled + height.cm_scaled + (1|pop) + (1|block)"
)

#run the models 
total_fit_models_CD_GD <- total_fit_models_CD_GD %>%
  mutate(glmmTMB = map(f, ~ glmmTMB(as.formula(.), 
                            family = nbinom1, ziformula=~1,
                            data = wl2_total_fitness_scaled)))

#show results:
total_fit_models_CD_GD %>% mutate(tidy=map(glmmTMB, tidy)) %>% unnest(tidy) %>%
  select(name, term:p.value) %>% 
  filter(str_detect(term, "scaled")) %>%
  drop_na(p.value) 

total_fit_models_CD_GD %>% 
  mutate(rsq=map(glmmTMB, r2),
         singularity=map(glmmTMB, check_singularity)) %>% 
  unnest(c(rsq, singularity)) %>% 
  unnest(rsq) %>% 
  select(-f:-glmmTMB)

##predictions for figures:
WY_Recent_mod <- total_fit_models_CD_GD %>% 
  filter(name=="2_WY_Recent") %>% 
  pull(glmmTMB) 
totalfitness_preds <- tibble(.pred=predict(WY_Recent_mod[[1]],
                                           newdata=pred_df_gowers,
                                           re.form= NA, type="response")) %>% 
  bind_cols(pred_df_gowers) 
totalfitness_preds$AvgGD_Recent_Wtr_Year_scaled <- as.vector(totalfitness_preds$AvgGD_Recent_Wtr_Year_scaled)
write_csv(totalfitness_preds, "../Processed.Data/TotalFitness_GDPreds_WYRecent.csv")
```

## Quasi-Poisson

### Basic Model
```{r}
library(lmerTest)

total_fit_poisson <- glmer(Total_Fitness ~ height.cm_scaled + (1|pop/mf) + (1|block), 
                              family = poisson,
                              data=wl2_total_fitness_scaled)
#results
summary(total_fit_poisson)
r2(total_fit_poisson) #marginal r2 = 0 

#diagnostics 
deviance(total_fit_poisson)/df.residual(total_fit_poisson) #check for overdispersion (caused by excess of 0s) - over 1 which is not good 
plot(total_fit_poisson, which = 1) #residuals plot (residuals vs predicted)
plot(predict(total_fit_poisson, type="response"), residuals(total_fit_poisson),
     main="poisson", xlab="Fitted", ylab="Residuals")  #same plot 
#doesn't look great
plot(residuals(total_fit_poisson)) #index plot 

## quasi poisson correction:
phi_function <- function(model) {
    phi <- sum(residuals(model, type="pearson")^2)/df.residual(model) 
    return(phi)
} #function for calculating the overdispersion paramter (for quasi-poisson)

##Can't directly use quasi families with glmer so have to use Ben Bolker's quasi-likelihood adjustment function:
#https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#fitting-models-with-overdispersion
#also got help from: https://stackoverflow.com/questions/68915173/how-do-i-fit-a-quasi-poisson-model-with-lme4-or-glmmtmb
quasi_table <- function(model,ctab=coef(summary(model))) {
    phi <- sum(residuals(model, type="pearson")^2)/df.residual(model) #overdispersion parameter
    qctab <- within(as.data.frame(ctab),
    {   `Std. Error` <- `Std. Error`*sqrt(phi)
        `z value` <- Estimate/`Std. Error`
        `Pr(>|z|)` <- 2*pnorm(abs(`z value`), lower.tail=FALSE)
    })
    return(qctab)
}
printCoefmat(quasi_table(total_fit_poisson),digits=2)
#As specified, the estimates are identical; the standard errors and p-values have been appropriately inflated, the z-values have been appropriately deflated.
```

### Recent water year Gowers model
```{r}
total_fit_poisson_wtryrGD <- glmer(Total_Fitness ~  AvgGD_Recent_Wtr_Year_scaled + Geographic_Dist_scaled + (1|pop/mf) + (1|block), 
                              family = poisson,
                              data=wl2_total_fitness_scaled)
summary(total_fit_poisson_wtryrGD)

##quasi correction:
printCoefmat(quasi_table(total_fit_poisson_wtryrGD),digits=2) #takes away significant recent water year result
```

## Comparing quasi-poisson to glmmTMB models 
Since the quasi part is based on the a p-value correction for overdispersion (doesn't change the model estimates), we can use the poisson model for all the diagnostics

### Residual Plot Comparisons 
```{r}
par(mfrow = c(2, 2))
plot(predict(total_fit_zero_negbinom1, type="response"), residuals(total_fit_zero_negbinom1),
     main="glmmTMB-negbinom", xlab="Fitted", ylab="Residuals",
     xlim=c(0, 60), ylim=c(-20, 120)) 
plot(predict(total_fit_zero_poisson, type="response"), residuals(total_fit_zero_poisson),
     main="glmmTMB-poisson", xlab="Fitted", ylab="Residuals",
     xlim=c(0, 60), ylim=c(-20, 120)) 
plot(predict(total_fit_poisson, type="response"), residuals(total_fit_poisson),
     main="poisson", xlab="Fitted", ylab="Residuals",
     xlim=c(0, 60), ylim=c(-20, 120)) 
```

### Predicted vs. Observed Comparisons 
```{r}
library(ggpubr) #ggarrange
negbinom_preds <- tibble(.pred=predict(total_fit_zero_negbinom1,
                                       newdata=wl2_total_fitness_scaled,
                                       type="response")) %>% 
  bind_cols(wl2_total_fitness_scaled) %>% 
  select(.pred, Total_Fitness)

negbinom_preds_fig <- negbinom_preds %>% 
  ggplot(aes(Total_Fitness, .pred)) +
  geom_point() +
  geom_abline(intercept=0, slope=1) +
  ylim(0, 70) +
  labs(x="Observed", y="Predicted", title="glmmTMB-negbinom")

zeriopoisson_preds <- tibble(.pred=predict(total_fit_zero_poisson,
                                       newdata=wl2_total_fitness_scaled,
                                       type="response")) %>% 
  bind_cols(wl2_total_fitness_scaled) %>% 
  select(.pred, Total_Fitness)

zeriopoisson_preds_fig <- zeriopoisson_preds %>% 
  ggplot(aes(Total_Fitness, .pred)) +
  geom_point() +
  geom_abline(intercept=0, slope=1) +
  ylim(0, 70) +
  labs(x="Observed", y="Predicted", title="glmmTMB-poisson")

poisson_preds <- tibble(.pred=predict(total_fit_poisson,
                                       newdata=wl2_total_fitness_scaled,
                                       type="response")) %>% 
  bind_cols(wl2_total_fitness_scaled) %>% 
  select(.pred, Total_Fitness)

poisson_preds_fig <- poisson_preds %>% 
  ggplot(aes(Total_Fitness, .pred)) +
  geom_point() +
  geom_abline(intercept=0, slope=1) +
  ylim(0, 70) +
  labs(x="Observed", y="Predicted", title="poisson")

ggarrange(negbinom_preds_fig, zeriopoisson_preds_fig, poisson_preds_fig,
                        ncol=2, nrow=2) 
```