---
title: "WklyClim_Mort"
author: "Brandie QC"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code for relating weekly climate to weekly mortality (Table 3)

## Load Libraries
```{r}
library(tidyverse)
library(zoo) #for rollmean function
library(tidymodels) #for model workflow 
tidymodels_prefer()
library(multilevelmod) #for using workflow with glmer 
library(broom.mixed) #for using glance with glmer 
```

## Read in the data
```{r}
wl2_surv <- read_csv("../Processed.Data/WL2_Mortality_2023.csv")
```

## As weeks
```{r}
wl2_surv_dates <- wl2_surv %>%  
  rename(parent.pop = pop) %>%
  mutate(parent.pop= str_replace(parent.pop, "Y08", "YO8")) %>% #fix cases where 0 used instead of O in pop name 
  mutate(parent.pop= str_replace(parent.pop, "Y04", "YO4")) %>% #fix cases where 0 used instead of O in pop name 
  filter(!is.na(parent.pop)) %>% 
  mutate(planting.date="7/19/23", #could try to make this more specific to when certain blocks were planted 
         last_fup_date=if_else(is.na(death.date), "10/27/23", death.date)) %>%  #need this to calculate survival times
  mutate(planting.date=mdy(planting.date), last_fup_date=mdy(last_fup_date)) %>% #convert to date objects
  mutate(os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1), #observed number of weeks
         status=if_else(is.na(death.date), 0, 1)) %>% #0=censured (alive in this case), 1=dead
  filter(os_weeks > 0) %>% #there is one case of a plant that was dead at planting, so just removed it since this is survival post- transplanting
  select(-survey.notes)
```

## Location Info

```{r}
pop_loc_WL2 <- read_csv("../Processed.Data/All_Clim_Dist.csv") %>% 
  select(parent.pop:Long)

#Merge
wl2_surv_dates_loc <- left_join(wl2_surv_dates, pop_loc_WL2) 
```

## Reformat to be one row per plant per week

```{r}
wl2_ind <- wl2_surv_dates %>% 
  complete(Genotype, last_fup_date) %>% #fill in missing weeks for each plant 
  group_by(Genotype) %>%
  mutate(across(c(-status, -last_fup_date, -os_weeks, -planting.date), 
                \(x) ifelse(all(is.na(unique(x))), NA,unique(na.omit(x))))) %>% #add block, loc, and pop/mf info back in 
  mutate(planting.date = unique(na.omit(planting.date)), #add planting date back in 
         os_weeks=as.duration(planting.date %--% last_fup_date) / dweeks(1), #add weeks back in 
           dead = ifelse((is.na(death.date) | mdy(death.date) > last_fup_date), 0, 1)) %>% #0 is alive, 1 is dead
  filter(cumsum(dead) < 2) %>% # you can't die more than once
  ungroup() %>%  
  left_join(pop_loc_WL2) #add location info back in 
```

## Read in soil temp

```{r}
temp <- read_csv("../Raw.Data/WL2_2022_2023_iButton_Data_Corrected.csv") %>%
  select(-`...3`) %>%
  mutate(Date_Time = mdy_hm(Date_Time)) %>%
  filter(Date_Time > ymd("2023-07-06"))
```

### Summarize
```{r}
temp_summary <- temp %>%
  group_by(Date_Time) %>% 
  summarise(AvgSoilTemp=mean(SoilTemp, na.rm=TRUE)) %>% #avg across beds 
  mutate(Date=as.Date(Date_Time)) %>%
  filter(Date != min(Date), Date != max(Date)) %>% # trim ragged ends
  group_by(Date) %>% #summarize hourly data by date
  summarize(
    min_temp_d = min(AvgSoilTemp), #min daily 
    max_temp_d = max(AvgSoilTemp), #max daily 
    mean_temp_d = mean(AvgSoilTemp), #mean daily 
    diurnal_temp_d = max_temp_d - min_temp_d
  ) %>%
  mutate(
    across(ends_with("temp_d"), \(x) rollmean(x, k = 7, align = "right", fill = NA), .names="{.col}1_7"), #summarize week prior to census 
    across(ends_with("temp_d"), \(x) rollmean(x, k = 13, align = "right", fill = NA), .names="{.col}1_13"), #summarize 2 weeks prior to census, used 13 to get the first survey date
    across(ends_with("1_7"), \(x) dplyr::lag(x, 6), .names = "{.col}_13") #summarize the first week of the above two weeks 
  ) %>%
  rename_with(\(x) str_replace(x, "1_7_13", "6_13"))
```

## Read in soil moisture

```{r}
moisture <- read_csv("../Raw.Data/WL2_2023_Bed_C_Soil_Moisture_Corrected.csv") %>%
  mutate(Date_Time = mdy_hm(Date_Time)) 
```

### Average across the ports

```{r}
moisture_summary <- moisture %>%
  mutate(Port_1=if_else(Port_1<0, 0, Port_1),
         Port_2=if_else(Port_2<0, 0, Port_2),
         Port_3=if_else(Port_3<0, 0, Port_3),
         Port_4=if_else(Port_4<0, 0, Port_4),
         Port_5=if_else(Port_5<0, 0, Port_5)) %>%  #convert negatives to zeros (negatives result from extremely dry soil --> air pockets)
  rowwise() %>%
  mutate(s_moisture = mean(c_across(-Date_Time)) ) %>%
  select(Date_Time, s_moisture) %>%
  mutate(Date=as.Date(Date_Time)) %>%
  group_by(Date) %>%
  summarize(
    mean_moisture_d = mean(s_moisture) #daily mean soil moisture 
  ) %>%
  mutate(
    s_moisture_1_7 = rollmean(mean_moisture_d, k = 7, align = "right", fill = "extend"), #summarize 1 week prior to census 
    s_moisture_1_14 = rollmean(mean_moisture_d, k = 13, align = "right", fill = "extend"), #summarize 2 weeks prior to census 
    s_moisture_7_14 = dplyr::lag(s_moisture_1_7, 7) #summarize the first week of the above two weeks  
  )
```

## Discrete intervals

Modify the input data to add some factor variables
```{r}
wl2_ind <- wl2_ind %>% mutate(
  os_weeks_f = factor(str_pad(round(os_weeks, 1), 2, pad = "0")), #round weeks column padding so it will sort correctly
  os_weeks_pop = factor(str_c(parent.pop,"_", os_weeks_f)) #make a column for pop_weeks
)
```

## Remove establishment period 
```{r}
wl2_ind_surv <- wl2_ind %>% filter(os_weeks > 3) %>%
  mutate(os_weeks_f = factor(os_weeks))
```

## Check basic model forms
```{r}
# Note: StdDev due to mf within parent.pop is essentially zero, so we omitted it
glmer.model <- 
  linear_reg() %>%
  set_engine("glmer", family = binomial)

surv_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, mf, block, os_weeks, os_weeks_f))

surv_fits_1 <- tibble(wflow=list(

  pop.block = {surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ (1|parent.pop) + (1|block))},
  
  pop.weeks_f.block ={surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks_f + (1|parent.pop) + (1|block))},

  pop.weeks.block ={surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop) + (1|block))},

  pop.weeks.quad.block ={surv_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + I(os_weeks^2) + (1|parent.pop) + (1|block))}
  ),
  name=names(wflow)) %>% 
  select(name,wflow)
```

```{r}
system.time(
  surv_fits_1 <- surv_fits_1 %>%
  mutate(fit = map(wflow, fit, data = wl2_ind_surv))
)
surv_fits_1 %>% pull(fit)
#convergence errors for pop.weeks_f.block and  pop.weeks.quad.block
```

## Add climate data

```{r}
wl2_ind_surv_climate <- wl2_ind_surv %>% 
  left_join(temp_summary, by = c("last_fup_date" = "Date")) %>%
  left_join(moisture_summary, by = c("last_fup_date" = "Date"))
```

## Temp analysis


```{r}
temp_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, block, os_weeks, contains("temp_d")  )) 

temp_fits <- tibble(wflow=list(

  # The best base model from earlier analyses:
  pop.weeks.block ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop) + (1|block))},
  
  min.temp.1.7 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ min_temp_d1_7 + os_weeks + (1|parent.pop) + (1|block))},
  
  mean.temp.1.7 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ mean_temp_d1_7 + os_weeks + (1|parent.pop) + (1|block))},
  
  max.temp.1.7 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ max_temp_d1_7 + os_weeks + (1|parent.pop) + (1|block))},
  
  diurnal.temp.1.7 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ diurnal_temp_d1_7 + os_weeks + (1|parent.pop) + (1|block))},
  
  min.temp.1.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ min_temp_d1_13 + os_weeks + (1|parent.pop) + (1|block))},
  
  mean.temp.1.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ mean_temp_d1_13 + os_weeks + (1|parent.pop) + (1|block))},
  
  max.temp.1.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ max_temp_d1_13 + os_weeks + (1|parent.pop) + (1|block))},
  
  diurnal.temp.1.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ diurnal_temp_d1_13 + os_weeks + (1|parent.pop) + (1|block))},
  
  min.temp.6.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ min_temp_d6_13 + os_weeks + (1|parent.pop) + (1|block))},
  
  mean.temp.6.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ mean_temp_d6_13 + os_weeks + (1|parent.pop) + (1|block))},
  
  max.temp.6.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ max_temp_d6_13 + os_weeks + (1|parent.pop) + (1|block))},
  
  diurnal.temp.6.13 ={temp_wflow %>% 
      add_model(glmer.model, formula = dead ~ diurnal_temp_d6_13 + os_weeks + (1|parent.pop) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
```

```{r}
system.time(
  temp_fits <- temp_fits %>%
  mutate(fit = map(wflow, fit, data = drop_na(wl2_ind_surv_climate, contains("temp_d")), .progress = TRUE))
)
```

```{r}
temp_fits %>% pull(fit)
```

```{r}
temp_fits %>% 
  mutate(glance=map(fit, glance)) %>% 
  unnest(glance) %>% 
  arrange(BIC) %>%
  select(-wflow:-fit, -sigma)
```

### Get p-values

```{r}
temp_fits %>%
  mutate(tidy=map(fit, tidy)) %>%
  select(name, tidy) %>%
  unnest(tidy) %>%
  filter(str_detect(term, "temp") | term=="os_weeks") %>%
  arrange(name) %>% 
  select(-effect, -group)
```

## Moisture analysis

```{r}
moisture_wflow <- workflow() %>%
  add_variables(outcomes = dead, predictors = c(parent.pop, block, os_weeks, contains("s_moisture") )) 

moisture_fits <- tibble(wflow=list(

  # The best base model from earlier analyses:
  pop.weeks.block ={moisture_wflow %>% 
      add_model(glmer.model, formula = dead ~ os_weeks + (1|parent.pop) + (1|block))},
  
    s_moisture1.7 ={moisture_wflow %>% 
      add_model(glmer.model, formula = dead ~ s_moisture_1_7 + os_weeks + (1|parent.pop) + (1|block))},
  
      s_moisture1.14 ={moisture_wflow %>% 
      add_model(glmer.model, formula = dead ~ s_moisture_1_14 + os_weeks + (1|parent.pop) + (1|block))},
      
          s_moisture7.14 ={moisture_wflow %>% 
      add_model(glmer.model, formula = dead ~ s_moisture_7_14 + os_weeks + (1|parent.pop) + (1|block))}
),
name=names(wflow)
) %>% 
  select(name,wflow)
```

```{r}
system.time(
  moisture_fits <- moisture_fits %>%
  mutate(fit = map(wflow, fit, data = drop_na(wl2_ind_surv_climate, contains("s_moisture")), .progress = TRUE))
)
```

```{r}
moisture_fits %>% pull(fit)
```

```{r}
moisture_fits %>% 
  mutate(glance=map(fit, glance)) %>% 
  unnest(glance) %>% 
  arrange(BIC) %>%
  select(-wflow:-fit, -sigma)
```

### Get p-values

```{r}
moisture_fits %>%
  mutate(tidy=map(fit, tidy)) %>%
  select(name, tidy) %>%
  unnest(tidy) %>%
  filter(str_detect(term, "moisture") | term=="os_weeks") %>% 
  arrange(name)
```